---
title: "07_Additional_Analysis"
author: "Reza Ghamsari"
date: "2025-05-19"
output: html_document
---
# Script Overview
This R Markdown document performs additional analyses on single-cell and single-nucleus RNA-seq data, focusing on:

Quality control visualization and assessment
Highly Variable Gene (HVG) identification and comparison between platforms
Gene length and GC content distribution analysis
Platform-specific technical bias characterization
```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now)
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

# Define Color Schemes and File Paths
```{r, colors}
# Custom color palette for consistent visualization
my_colors <- unique(c(
  "#802268FF", "#006400", "#525ecc99", "#1f0fdf99", "#480607", "#D55E00", 
  "#164194FF", "#7E6148FF", "#ADB6B6FF", "#52a7cc99", "#F0E685FF", "#cc52c099", 
  "#6acc5299", "#DC143C", "#802268FF", "#ffbb78", "#3C548899", "#189b3699", 
  "#cc9b5299", "#1f0fdf99", "#ccebc5", "#42857599", "#868686FF", "#006400", 
  "#0A47FFFF", "#3B1B53FF", "#749B58FF", "#ccc05299", "#cc765299", "#1B1919FF", 
  "#00D68FFF", "#14FFB1FF", "#3C5488FF", "#8F7700FF", "#164194FF", "#0094CDFF", 
  "#FFDAB9", "#FF00FF", "#00FFFF", "green", "blue", "#480607"
))

# Define the directory to save the PDF files
rds_path_vst <- "/vast/projects/aml_multiome/Sn_vs_SC/rds_path_vst/"
fig_path_stnx <- "/home/users/allstaff/ghamsari.r/seurat_figures/Sn_vs_SC_V4/"

```



# Loading Packages
```{r, pkgs}
script_title <- "06_02_EdgeR_pseudo_bulk_Analysis"
pkgs <- c("Signac", "Seurat", "BiocParallel", "DESeq2", "gridExtra", "parallel", "edgeR", "RColorBrewer", "patchwork", 
"limma", "tibble", "openxlsx", "ggplot2", "reshape2", "dplyr", "grid", "tidyverse", "htmlwidgets", "cowplot", "stringr", 
"tidyr", "pheatmap", "ggrepel" , "future" , "future.apply", "UpSetR" , "introdataviz")

# Load all packages silently
invisible(lapply(pkgs, library, character.only = TRUE))

# Save session information for reproducibility
sink(file.path(fig_path_stnx, paste0(script_title, "_sessionInfo.txt")))
sessionInfo()
sink()
```




# Reading data
```{r, reading objects}
merged_obj <- readRDS(paste0(rds_path_vst, "05_Merge_object_cell_types_annotated_clustered_list_22_samples_11sn_11sc.rds"))  #36601 features across 172055 samples
merged_obj_v0 <- merged_obj
merged_obj_list<- SplitObject(merged_obj, split.by = "sample_id")
```
# Quality Control Analysis

Scatter Plot Visualization by Cell Type: Examine QC metrics patterns across different cell types to identify platform-specific biases, particularly in erythroid populations 

```{r, scatter plot and MAD }
###made this plot to re-check the plot 04_scatter_plots_visualisation_all_cells_before_subsetting_MAD_6 and see which cells are responsible for seeing two trend in scRNAseq.
scatter_features <- list(
  c("nCount_RNA", "nFeature_RNA"),     # Total counts vs unique genes
  c("percent.mt", "percent.malat1")    # Mitochondrial vs MALAT1 percentage
)


 
# Function to create scatter plots colored by cell type
create_plot <- function(features, sobj, sample_id) {
  # Extract the actual feature values
  x_feature <- features[1]
  y_feature <- features[2]
  # Prepare data frame for plotting
  plot_df <- data.frame(
    x = sobj@meta.data[[x_feature]],
    y = sobj@meta.data[[y_feature]],
    cell_type = as.factor(sobj@meta.data[["cell_type_fine"]]),
    Donor = as.factor(sobj@meta.data[["sample_id"]]),
    colors_cell_type = sobj@meta.data[["colors_cell_type_fine"]]
  )
  
  # Build color mapping from cell type to its color

  color_mapping <- setNames(levels(plot_df$colors_cell_type ),levels(plot_df$cell_type))
  
  # Generate scatter plot 
  scatter_plot <- ggplot(plot_df, aes(x = x, y = y, color = cell_type)) +
    geom_point(alpha = 0.8, size = 0.9) +
    scale_color_manual(
      values = color_mapping,
      guide = guide_legend(override.aes = list(size = 9, alpha = 1))
    ) +
    labs(
      title = paste(sample_id),
      x = x_feature,
      y = y_feature,
      color = "Cell Type"
    ) +
    theme_minimal() +
    theme(panel.background = element_rect(fill = "white"),
      panel.grid = element_blank(),
      plot.title = element_text(hjust = 0.5)
    )

  return(scatter_plot)
}




# Generate comprehensive QC plots for all samples
pdf(file = paste0(fig_path_stnx, "07_scatter_qc_plot_color_by_cell_type_fine_per_sample_split.pdf"), width = 14, height = 7)

# Loop over all samples
for (sample_id in names(merged_obj_list)) {
  
  # Extract the current sample object
  sobj <- merged_obj_list[[sample_id]]
  
  # Create a list to store the plots for the current sample
  plots <- list()
  
  # Loop over all feature pairs
  for (features in scatter_features) {
    
    # Create the plot for the current feature pair
    plot <- create_plot(features, sobj, sample_id)
    
    # Add the plot to the list
    plots[[length(plots) + 1]] <- plot
  }
  
  # Arrange and display the plots on one page
  grid.arrange(grobs = plots, ncol = 2)
}

dev.off()
#!!scRNA-seq shows two distinct trends in erythroblast populations
```
# Highly Variable Gene (HVG) Analysis
HVG Identification Using Multiple Methods
Purpose: Compare HVG identification between SCTransform and variance stabilizing transformation (VST) methods across methodologies (platforms)
 to assess method robustness and platform effects.


```{r, clustering}
# To identify the number of available cores
num_cores <- parallelly::availableCores() #56
plan(multisession, workers = num_cores - 4)

#============================================== HVGs Sctransform ===============================================================
processSeurat <- function(x) {
  x <- SCTransform (x, vst.flavor = "v2", verbose = FALSE,new.assay.name = "SCT_v2", seed.use = 1448145, variable.features.n = 3000)
  hvgs <-   x@assays$SCT_v2@var.features
    return(hvgs)
}
# Generate HVGs for individual samples using SCTransform
HVG_list_sct <- future_lapply(merged_obj_list, processSeurat, future.seed = TRUE)
saveRDS(HVG_list_sct, paste0(rds_path_vst, "07_HVG_list_sct_v2_3000.rds"))
# HVG_list_sct <- readRDS(paste0(rds_path_vst, "07_HVG_list_sct_v2_3000.rds"))
#========================================
# Generate HVGs for merged object using SCTransform
merged_obj_sct <- SCTransform (merged_obj, vst.flavor = "v2", verbose = FALSE,new.assay.name = "SCT_v2", seed.use = 1448145, variable.features.n = 3000)
merge_hvgs_sct <-   merged_obj_sct@assays$SCT_v2@var.features
saveRDS(merge_hvgs_sct, paste0(rds_path_vst, "07_HVG_merged_obj_sct_v2_3000.rds"))
# merge_hvgs_sct <- readRDS(paste0(rds_path_vst, "07_HVG_merged_obj_sct_v2_3000.rds"))

#==================================================HVGs VST seurat Default==============================================
processSeurat <- function(x) {
    x <- NormalizeData(x, assay="RNA")
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures= 3000, assay="RNA")
    hvg_vst <- VariableFeatures(x[["RNA"]])
    return(hvg_vst)
}
# Generate HVGs for individual samples using VST
HVG_list_vst <- future_lapply(merged_obj_list, processSeurat, future.seed = TRUE)
saveRDS(HVG_list_vst, paste0(rds_path_vst, "07_HVG_list_vst_3000.rds"))
# HVG_list_vst <- readRDS(paste0(rds_path_vst, "07_HVG_list_vst_3000.rds"))
# Generate HVGs for merged object using VST
merged_obj <- merged_obj_v0 #36601 features across 172055 samples
merged_obj <- NormalizeData(merged_obj, assay="RNA")
merged_obj<- FindVariableFeatures(merged_obj, selection.method = "vst", nfeatures= 3000, assay="RNA")
merge_hvgs_vst <- VariableFeatures(merged_obj[["RNA"]])
saveRDS(merge_hvgs_vst, paste0(rds_path_vst, "07_HVG_merged_obj_vst_3000.rds"))
# merge_hvgs_vst <- readRDS(paste0(rds_path_vst, "07_HVG_merged_obj_vst_3000.rds"))

```
## Intra-Platform HVG Consistency Analysis
Assess the robustness of HVG identification within each platform using Jaccard similarity indices.

```{r}
# pdf(file = paste0(fig_path_stnx, "07_Jaccard_split_violin_HVGs_3000_SCT.pdf"), width = 12, height = 16)
# HVG_list <- readRDS(paste0(rds_path_vst, "07_HVG_list_sct_v2_3000.rds"))
# pdf(file = paste0(fig_path_stnx, "07_Jaccard_split_violin_HVGs_3000_VST.pdf"), width = 12, height = 16)
# HVG_list <- readRDS(paste0(rds_path_vst, "07_HVG_list_vst_3000.rds"))

# Separate samples by platform
sc_HVGs <- HVG_list[grep("_cell", names(HVG_list))]
sn_HVGs <- HVG_list[grep("_Nuclei", names(HVG_list))]

# Function to calculate pairwise Jaccard similarities within a group
get_jaccard_distribution <- function(hvg_list) {
  n <- length(hvg_list)
  jaccard_values <- c()
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      jaccard <- length(intersect(hvg_list[[i]], hvg_list[[j]])) /
                 length(union(hvg_list[[i]], hvg_list[[j]]))
      jaccard_values <- c(jaccard_values, jaccard)
    }
  }
  return(jaccard_values)
}
# Calculate Jaccard distributions for each platform
sc_jaccards <- get_jaccard_distribution(sc_HVGs)
sn_jaccards <- get_jaccard_distribution(sn_HVGs)


# Prepare data for visualization
boxplot_data <- list(scRNAseq = sc_jaccards, snRNAseq = sn_jaccards)


# Convert list to dataframe for ggplot

df <- data.frame(
  Jaccard = c(boxplot_data$scRNAseq, boxplot_data$snRNAseq),
  Platform = rep(c("scRNAseq", "snRNAseq"),
                 times = c(length(boxplot_data$scRNAseq), length(boxplot_data$snRNAseq)))
)

ggplot(df, aes(x = Platform, y = Jaccard, fill = Platform)) +
  geom_violin(trim = FALSE, alpha = 0.3, color = NA) +

  geom_jitter(width = 0.05, size = 1.5, alpha = 0.7) + 
  stat_summary(fun.data = mean_se, geom = "crossbar", show.legend = TRUE,
               width = 0.3, color = "black", fatten = 2) +  # mean ± SE bars
  scale_fill_manual(values = c("scRNAseq" = "#6699CC", "snRNAseq" = "#CC6677")) +
  theme_minimal() +
  labs(title = "Intra-platform HVG Consistency",
       y = "Jaccard Index (HVG Overlap)",
       x = "") +
theme(
  axis.line = element_line(color = "black"),  # add axis lines
  axis.ticks = element_line(color = "black"), # add ticks
    legend.position = "none",
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    panel.grid = element_blank(),        # remove all grid lines
    panel.border = element_blank()      # remove panel border if any
  )



####################################### split violin plot
# Add fixed x for violin center
df <- df %>%
  mutate(
    x = 0,
    # Custom x offsets: left half for scRNAseq, right half for snRNAseq
    x_jitter = case_when(
      Platform == "scRNAseq" ~ -0.03,
      Platform == "snRNAseq" ~ 0.03
    ),
    x_summary = case_when(
      Platform == "scRNAseq" ~ -0.02,
      Platform == "snRNAseq" ~ 0.02
    )
  )

ggplot(df, aes(x = x, y = Jaccard, fill = Platform)) +
  geom_split_violin(alpha = 0.3, trim = FALSE, color = "transparent", width = 0.8, show.legend = TRUE) +
  
  # Points jittered and nudged left/right per Platform
  geom_point(aes(x = x_jitter, color = Platform), size = 0.1, alpha = 0.7, position = position_jitter(width = 0)) +
    geom_boxplot(aes(x = x_jitter, fill = Platform), width = 0.025, alpha = 0.4, size = 0.05, fatten = NULL, show.legend = TRUE, outlier.size = 0.5, color = "white",
                   outlier.color = "gray", outlier.alpha = 0.6) +

  stat_summary(aes(x = x_jitter, color = Platform), fun.data = mean_se, geom = "pointrange",
               width = 0.9,  fatten = 5,
               position = position_identity()) +
  
scale_fill_manual(values = c("scRNAseq" = "#6699CC", "snRNAseq" = "#CC6677")) +
  scale_color_manual(values = c("scRNAseq" = "#00468B", "snRNAseq" = "#8B0046")) +  
    scale_y_continuous(
    breaks = seq(0, 1, by = 0.1),
    limits = c(0, 1)
  ) +
  
  labs(
    title = "Intra-platform HVG Consistency",
    x = "", 
    y = "Jaccard Index (HVG Overlap)"
  ) +
  theme_minimal() +
  theme(
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14),
    legend.position = "top",
    panel.grid = element_blank(),
    panel.border = element_blank()
  )

dev.off()
```
## HVG Similarity Heatmap Analysis
Purpose: Visualize sample-to-sample HVG similarities using hierarchical clustering to identify platform and batch effects.

#jaccard heatmap
```{r, fig.width=10}
# Load HVG data and add merged object HVGs
#========================= variance stabilizing transform
HVG_list <- readRDS(paste0(rds_path_vst, "07_HVG_list_vst_3000.rds"))
merge_hvgs_vst <- readRDS(paste0(rds_path_vst, "07_HVG_merged_obj_vst_3000.rds"))
HVG_list[["ALL_ALL_Merged"]] <- merge_hvgs_vst
pdf_file <- paste0(fig_path_stnx, "07_Jaccard_heatmap_HVGs_3000_vst_v_final.pdf")
main_title <-  "Jaccard (ward.D2) of HVGs VST"


#=========================. ScTransform
# HVG_list <- readRDS(paste0(rds_path_vst, "07_HVG_list_sct_v2_3000.rds"))
# merge_hvgs_sct <- readRDS(paste0(rds_path_vst, "07_HVG_merged_obj_sct_v2_3000.rds"))
# HVG_list[["ALL_ALL_Merged"]] <- merge_hvgs_sct
# pdf_file <- paste0(fig_path_stnx, "07_Jaccard_heatmap_HVGs_3000_SCT_v_final.pdf")
# main_title <-  "Jaccard (ward.D2) of HVGs SCT"


#========================= Function to calculate full Jaccard similarity matrix
# All pairwise Jaccard within a group
names(HVG_list)

get_jaccard_matrix <- function(hvg_list) {
  n <- length(hvg_list)
  mat <- matrix(NA, n, n, dimnames = list(names(hvg_list), names(hvg_list)))
 for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      mat[i, j] <- length(intersect(hvg_list[[i]], hvg_list[[j]])) / 
                   length(union(hvg_list[[i]], hvg_list[[j]]))
    }
 }

  return(mat)
}

# Calculate Jaccard similarity matrix
mtx_jaccards <- get_jaccard_matrix(HVG_list)

# Convert upper triangle to full symmetric matrix
mtx_jaccards[lower.tri(mtx_jaccards)] <- t(mtx_jaccards)[lower.tri(mtx_jaccards)]

# Define annotation colors
my_annotation_colors <- list(
  Method = c(cell = "#6699CC", Nuclei = "#CC6677", Merged ="white"),
  Site = c(S1 = "#661100", S2 = "#888888", S3 = "#008B45", S4 = "#D55E00", ALL= "white" ),
  Donor = c(D1 = "#00A087FF", D2 = "#4A0F5FFF", D3 = "#00468B", D4 = "#BB0021FF",
            D5 = "#006400", D6 = "#08306B", D7 = "#ffbb78", D8 = "#8da0cb",
            D9 = "#e78ac3", ALL= "white")
)

# Create sample metadata for annotations
meta <- data.frame(Sample = rownames(mtx_jaccards)) %>%
  separate(Sample, into = c("Site", "Donor", "Method"), sep = "_",  remove = FALSE)

# Use as annotation row/column with reordered columns
annotation_col <- meta %>%
  select(Sample, Method) %>%  # re-order columns
  column_to_rownames("Sample")

annotation_row <- meta %>%
  select(Sample, Site, Donor) %>%  # re-order columns
  column_to_rownames("Sample")

# Perform clustering using Ward.D2 method on MDS-embedded distances
dist_jaccard <- as.dist(1 - mtx_jaccards)

# Embed into Euclidean space
mds_coords <- cmdscale(dist_jaccard, k = 16) #Classical Multidimensional Scaling (MDS) to embed the samples into a Euclidean space to approximate non-Euclidean distances (like Jaccard)

# Compute Euclidean distances on embedded data
dist_euc <- dist(mds_coords)

# Ward.D2 clustering on Euclidean distances
hc_ward <- hclust(dist_euc, method = "ward.D2")

# Generate heatmap
pdf(pdf_file, width = 7, height = 6)
color_vector <- c("white", "#737373")
# Plot
fontsize = 9

# Use hc_ward in heatmap
pheatmap(mtx_jaccards,
         cluster_rows = hc_ward,
         cluster_cols = hc_ward,
         clustering_distance_rows = NULL,
         clustering_distance_cols = NULL,
         clustering_method = NULL,
         annotation_col = annotation_col,
         annotation_row = annotation_row,
         annotation_colors = my_annotation_colors,
         fontsize_row = fontsize,
         fontsize_col = fontsize,
         main = main_title,
         color = colorRampPalette(color_vector)(400),
         treeheight_col = 30,
         treeheight_row = 0)

dev.off()


##uncomment for scTransform

# pheatmap(mtx_jaccards,
#          clustering_distance_rows = dist_jaccard,
# clustering_distance_cols = dist_jaccard,
#          clustering_method = "average",
#           cluster_rows = TRUE,
#          cluster_cols = TRUE,
#          annotation_col = annotation_col,
#          annotation_row = annotation_row,
#         annotation_colors = my_annotation_colors,
#          main = "Jaccard Similarity of HVGs (variance stabilizing transform/average clustering)",
#    # main = "Jaccard Similarity of HVGs (scTransform/average clustering)",
# 
#         color = colorRampPalette(color_vector))(100),
#   treeheight_row = 0,
#          treeheight_col = 0
# )


```

# Gene Length Distribution Analysis
Examine how gene length affects count distribution between scRNA-seq and snRNA-seq platforms.
```{r, final}
merged_obj <- merged_obj_v0 #36601 features across 172055 samples
merged_obj_list<- SplitObject(merged_obj, split.by = "sample_id")

############ Define the gene length categories

for (obj in names(merged_obj_list)){
  sobj <- merged_obj_list[[obj]]
  # Step 1: Get the gene row sums across all barcodes (total count per gene)
  count_matrix <- sobj@assays$RNA$counts
  nCount_per_gene <- rowSums(count_matrix)
 if(identical(names(nCount_per_gene), rownames(sobj))) {#TRUE
 sobj@assays$RNA@meta.data$nCount_per_gene <- nCount_per_gene}
  merged_obj_list[[obj]] <- sobj
}
# Analyze gene length distribution across samples
all_counts <- list()

for (obj in names(merged_obj_list)) {
  sobj <- merged_obj_list[[obj]]
  platform <- unique(sobj$Platform)
  Site <- unique(sobj$Site)
  Donor <- unique(sobj$Donor)
  Donor_Site <- unique(sobj$Donor_Site)
  gene_df <- sobj@assays$RNA@meta.data
  gene_df <- gene_df[!is.na(gene_df$gene_length), ]

  gene_df <- gene_df %>%
  arrange(gene_length)
  dens <- density(gene_df$gene_length, na.rm = TRUE)
# # Plot it to inspect
#  plot(dens, main = "Density Estimate of Gene Lengths")
# abline(v = median(gene_df$gene_length), col = "red", lty = 2)  # Add the raw data median

  # Create gene length bins (50 quantile-based bins)
  gene_df$length_bin <- cut(
    gene_df$gene_length,
    breaks = quantile(gene_df$gene_length, probs = seq(0, 1, length.out = 50)),
    include.lowest = TRUE
  )
# Calculate counts per bin
  counts_by_bin <- gene_df %>%
  group_by(length_bin) %>%
  summarise(
    total_counts = sum(nCount_per_gene),
    gene_number = n()
  ) %>%
  mutate(
    platform = platform,
    Donor = Donor,
    Site = Site,
    Donor_Site = Donor_Site,
    Site_platform = paste0(Site, "_", platform),
    Sample_id = obj
  )
table(counts_by_bin$gene_number)
  all_counts[[obj]] <- counts_by_bin
}

# Combine all into one dataframe
combined_df <- bind_rows(all_counts)


# Extract bin midpoints for plotting
scientific_num_pattern <- "[0-9\\.eE\\+\\-]+"
combined_df <- combined_df %>%
  mutate(
    bin_start = as.numeric(str_extract(length_bin, paste0("(?<=\\[|\\()", scientific_num_pattern))),
    bin_end = as.numeric(str_extract(length_bin, paste0("(?<=,)", scientific_num_pattern, "(?=\\]|\\))"))),
    bin_mid = (bin_start + bin_end) / 2
  )

combined_df$bin_mid_factor <- factor(combined_df$bin_mid)

combined_df <- combined_df %>%
  group_by(Sample_id) %>%
  mutate(
    total_counts_per_sample = sum(total_counts),  # Total counts for the entire sample
    # Normalize within each sample and calculate percentage of total counts per bin
    percentage_bin_count = (total_counts / total_counts_per_sample) * 100
  ) %>%
  ungroup()

# Create visualization
pdf(file = paste0(fig_path_stnx, "07_Distribution_gene_length_v2.pdf"), width = 8, height = 5)
ggplot(combined_df, aes(x = bin_mid_factor, y = percentage_bin_count, color = platform, group = Sample_id)) +
  geom_line(size = 0.5, alpha = 0.5) +
  labs(
    title = "Percentage of Counts per Gene-Length-Bin per sample Across 11 Paired Samples \n 50 bins ~720 genes per bin",
    x = "Gene Length Bin Midpoint (bp)",
    y = "Counts per Bin / per sample *100",
    color = "Sample"
  ) +
  scale_color_manual(values = c("scRNA" = "#00468B", "snRNA" = "#8B0046")) +
  scale_x_discrete(breaks = levels(combined_df$bin_mid_factor)[seq(1, length(levels(combined_df$bin_mid_factor)), by = 2)])+
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.line = element_line(color = "black"),
    plot.title = element_text(size = 8)
  )

dev.off()
```

## GC Content Distribution Analysis
Analyze GC content bias between platforms, as GC-rich regions may be differentially accessible to nuclear vs cellular extraction methods.

```{r, final}
merged_obj <- merged_obj_v0 #36601 features across 172055 samples
merged_obj_list<- SplitObject(merged_obj, split.by = "sample_id")

############ Define the gene GC-Percentage categories


for (obj in names(merged_obj_list)){
  sobj <- merged_obj_list[[obj]]
  # Step 1: Get the gene row sums across all barcodes (total count per gene)
  count_matrix <- sobj@assays$RNA$counts
  nCount_per_gene <- rowSums(count_matrix)
 if(identical(names(nCount_per_gene), rownames(sobj))) {#TRUE
 sobj@assays$RNA@meta.data$nCount_per_gene <- nCount_per_gene}
  merged_obj_list[[obj]] <- sobj
}

all_counts <- list()

for (obj in names(merged_obj_list)) {
  sobj <- merged_obj_list[[obj]]
  platform <- unique(sobj$Platform)
  Site <- unique(sobj$Site)
  Donor <- unique(sobj$Donor)
  Donor_Site <- unique(sobj$Donor_Site)
  gene_df <- sobj@assays$RNA@meta.data
  gene_df <- gene_df[!is.na(gene_df$percentage_gene_gc_content), ]
  gene_df <- gene_df %>%
  arrange(percentage_gene_gc_content)
  dens <- density(gene_df$percentage_gene_gc_content, na.rm = TRUE)
# 
# # Plot it to inspect
#  plot(dens, main = "Density Estimate of Gene GC percentage")
# abline(v = median(gene_df$percentage_gene_gc_content), col = "red", lty = 2)  # Add the raw data median

  # Create bins - use fewer bins for clarity, e.g. 100
  gene_df$GC_bin <- cut(
    gene_df$percentage_gene_gc_content,
    breaks = quantile(gene_df$percentage_gene_gc_content, probs = seq(0, 1, length.out = 100)),
    include.lowest = TRUE
  )

  counts_by_bin <- gene_df %>%
  group_by(GC_bin) %>%
  summarise(
    total_counts = sum(nCount_per_gene),
    gene_number = n()
  ) %>%
  mutate(
    platform = platform,
    Donor = Donor,
    Site = Site,
    Donor_Site = Donor_Site,
    Site_platform = paste0(Site, "_", platform),
    Sample_id = obj
  )
  all_counts[[obj]] <- counts_by_bin
}

# Combine all into one dataframe
combined_df <- bind_rows(all_counts)



scientific_num_pattern <- "[0-9\\.eE\\+\\-]+"
combined_df <- combined_df %>%
  mutate(
    bin_start = as.numeric(str_extract(GC_bin, paste0("(?<=\\[|\\()", scientific_num_pattern))),
    bin_end = as.numeric(str_extract(GC_bin, paste0("(?<=,)", scientific_num_pattern, "(?=\\]|\\))"))),
    bin_mid = (bin_start + bin_end) / 2
  )

combined_df$bin_mid_factor <- factor(combined_df$bin_mid)

combined_df <- combined_df %>%
  group_by(Sample_id) %>%
  mutate(
    total_counts_per_sample = sum(total_counts),  # Total counts for the entire sample
    # Normalize within each sample and calculate percentage of total counts per bin
    percentage_bin_count = (total_counts / total_counts_per_sample) * 100
  ) %>%
  ungroup()



pdf(file = paste0(fig_path_stnx, "07_Distribution_percentage_gene_gc_content_v2.pdf"), width = 8, height = 5)
ggplot(combined_df, aes(x = bin_mid_factor, y = percentage_bin_count, color = platform, group = Sample_id)) +
  geom_line(size = 0.5, alpha = 0.5) +
  labs(
    title = "Percentage of Counts per GC-Percentage-Bin per sample Across 11 Paired Samples \n 50 bins ~720 genes per bin",
    x = "Gene GC-Percentage Bin Midpoint (bp)",
    y = "Percentage Counts per Bin per sample",
    color = "Sample"
  ) +
  scale_color_manual(values = c("scRNA" = "#00468B", "snRNA" = "#8B0046")) +
  scale_x_discrete(breaks = levels(combined_df$bin_mid_factor)[seq(1, length(levels(combined_df$bin_mid_factor)), by = 4)])+
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.line = element_line(color = "black")
  )


dev.off()

```

### Finding the genes with high percentage in length bins
```{r, }
obj <- "S1_D1_cell"

for (obj in names(merged_obj_list_Sub)){
 sobj <- merged_obj_list_Sub[[obj]]
  # Step 1: Get the gene row sums across all barcodes (total count per gene)
  count_matrix <- sobj@assays$RNA$counts
  nCount_per_gene <- rowSums(count_matrix)
  lib_size <- sum(nCount_per_gene)
  nCount_per_gene_cpm <- nCount_per_gene/lib_size*1000000
 if(identical(names(nCount_per_gene), rownames(sobj))) {#TRUE
 sobj@assays$RNA@meta.data$nCount_per_gene <- nCount_per_gene
 sobj@assays$RNA@meta.data$nCount_per_gene_cpm <- nCount_per_gene_cpm}
  platform <- unique(sobj$Platform)
  Site <- unique(sobj$Site)
  Donor <- unique(sobj$Donor)
  Donor_Site <- unique(sobj$Donor_Site)
  gene_df <- sobj@assays$RNA@meta.data
  gene_df <- gene_df[!is.na(gene_df$gene_length), ]
  gene_df <- gene_df %>%
  arrange(gene_length)
  # Create bins - use fewer bins for clarity, e.g. 100
  gene_df$length_bin <- cut(
    gene_df$gene_length,
    breaks = quantile(gene_df$gene_length, probs = seq(0, 1, length.out = 50)),
    include.lowest = TRUE
  )

  counts_by_bin <- gene_df %>%
  group_by(length_bin) %>%
  summarise(
    total_counts = sum(nCount_per_gene),
    gene_number = n()
  ) %>%
  mutate(
    platform = platform,
    Donor = Donor,
    Site = Site,
    Donor_Site = Donor_Site,
    Site_platform = paste0(Site, "_", platform),
    Sample_id = obj
  )


# Combine all into one dataframe
combined_df <- counts_by_bin

combined_df <- combined_df %>%
  mutate(
    total_counts_per_sample = sum(total_counts),  # Total counts for the entire sample
    # Normalize within each sample and calculate percentage of total counts per bin
    percentage_bin_count = (total_counts / total_counts_per_sample) * 100
  ) 
# Sort the combined_df by percentage_bin_count in descending order
arranged_combined_df_descending <- combined_df %>%
  arrange(desc(percentage_bin_count))

# View the result
head(arranged_combined_df_descending)

# Find the bin with the highest percentage (from previous steps)
# highest_percentage_bin <- combined_df %>%
#   filter(percentage_bin_count == max(percentage_bin_count, na.rm = TRUE))
highest_percentage_bin <- combined_df %>%
  filter(percentage_bin_count >4 )
highest_percentage_bin$percentage_bin_count
# Get the length_bin of the bin with the highest percentage
highest_bin <- highest_percentage_bin$length_bin


for (bin in highest_bin) {
# Filter gene_df to find genes in the bin with the highest percentage
genes_in_highest_bin <- gene_df %>%
  filter(length_bin == bin)

arranged_genes_in_highest_bin <- genes_in_highest_bin %>%
  arrange(desc(nCount_per_gene))
print(paste0("Top genes in bin ", bin, " for sample ", obj, ": ", 
             paste(arranged_genes_in_highest_bin$hgnc_symbol[1:3], collapse = ", ")," CPM= ", paste(arranged_genes_in_highest_bin$nCount_per_gene_cpm[1:3], collapse = ", ")))

}}

# Filter gene_df to find the gene with hgnc_symbol "HBA2"
hba2_gene <- gene_df %>%
  filter(hgnc_symbol == "HBA2")
```
### Removing peaks in the first plot and re-visualization
#### Gene length visualization after filtering some genes like HBs and RPs

```{r, length}
merged_obj <- merged_obj_v0 #36601 features across 172055 samples
all_genes_name <- rownames(merged_obj@assays$RNA)
removing_genes_patterns <- "^HB[^(P)]|^RP[SL]|^MT-|MALAT1"
# removing_genes_patterns <- "^IGHV|^IGKV|^IGLV|^TRAV|^TRBV|^TRDV|^TRGV|^HB[^(P)]|MALAT1|^RP[SL]|^IG[HKL][A-Z0-9-]*|^MT-"
removing_genes <- grep(removing_genes_patterns, all_genes_name, value = TRUE)
length(removing_genes) #128
residual_genes <- all_genes_name[!all_genes_name %in% removing_genes]
length(residual_genes) #36473

merged_obj_sub <- subset(merged_obj, features = residual_genes) #34920 features across 172055 samples within 1 assay

merged_obj_list_Sub<- SplitObject(merged_obj_sub, split.by = "sample_id")

# #############
############ Define the gene length categories
#############

all_counts <- list()
for (obj in names(merged_obj_list_Sub)){
  sobj <- merged_obj_list_Sub[[obj]]
  # Step 1: Get the gene row sums across all barcodes (total count per gene)
  count_matrix <- sobj@assays$RNA$counts
  nCount_per_gene <- rowSums(count_matrix)
 if(identical(names(nCount_per_gene), rownames(sobj))) {#TRUE
 sobj@assays$RNA@meta.data$nCount_per_gene <- nCount_per_gene}
  platform <- unique(sobj$Platform)
  Site <- unique(sobj$Site)
  Donor <- unique(sobj$Donor)
  Donor_Site <- unique(sobj$Donor_Site)
  gene_df <- sobj@assays$RNA@meta.data
  gene_df <- gene_df[!is.na(gene_df$gene_length), ]

 gene_df <- gene_df %>%
  arrange(gene_length)
  # Create bins - use fewer bins for clarity, e.g. 100
  gene_df$length_bin <- cut(
    gene_df$gene_length,
    breaks = quantile(gene_df$gene_length, probs = seq(0, 1, length.out = 50)),
    include.lowest = TRUE
  )

  counts_by_bin <- gene_df %>%
  group_by(length_bin) %>%
  summarise(
    total_counts = sum(nCount_per_gene),
    gene_number = n()
  ) %>%
  mutate(
    platform = platform,
    Donor = Donor,
    Site = Site,
    Donor_Site = Donor_Site,
    Site_platform = paste0(Site, "_", platform),
    Sample_id = obj
  )

  all_counts[[obj]] <- counts_by_bin
}

# Combine all into one dataframe
combined_df <- bind_rows(all_counts)



scientific_num_pattern <- "[0-9\\.eE\\+\\-]+"
combined_df <- combined_df %>%
  mutate(
    bin_start = as.numeric(str_extract(length_bin, paste0("(?<=\\[|\\()", scientific_num_pattern))),
    bin_end = as.numeric(str_extract(length_bin, paste0("(?<=,)", scientific_num_pattern, "(?=\\]|\\))"))),
    bin_mid = (bin_start + bin_end) / 2
  )

combined_df$bin_mid_factor <- factor(combined_df$bin_mid)

combined_df <- combined_df %>%
  group_by(Sample_id) %>%
  mutate(
    total_counts_per_sample = sum(total_counts),  # Total counts for the entire sample
    # Normalize within each sample and calculate percentage of total counts per bin
    percentage_bin_count = (total_counts / total_counts_per_sample) * 100
  ) %>%
  ungroup()



pdf(file = paste0(fig_path_stnx, "07_Distribution_gene_length_HB_RB_Removed_v3.pdf"), width = 8, height = 5)
ggplot(combined_df, aes(x = bin_mid_factor, y = percentage_bin_count, color = platform, group = Sample_id)) +
  geom_line(size = 0.5, alpha = 0.5) +
  labs(
    title = "Percentage of Counts per Gene Length Bin Across 11 Paired Samples \n MTs, RBs, HBs and Malat1 removed \n 50 bin ~720 genes per bin",
    x = "Gene Length Bin Midpoint (bp)",
    y = "Counts per Bin / per sample *100",
    color = "Sample"
  ) +
  scale_color_manual(values = c("scRNA" = "#00468B", "snRNA" = "#8B0046")) +
  scale_x_discrete(breaks = levels(combined_df$bin_mid_factor)[seq(1, length(levels(combined_df$bin_mid_factor)), by = 2)])+
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.line = element_line(color = "black"),
    plot.title = element_text(size = 8) 
  )


dev.off()
```
#### Gene GC  re-visualization after filtering some genes like HBs and RPs

```{r, GC}

#############
############ Define the gene GC-Percentage categories
#############

for (obj in names(merged_obj_list_Sub)){
  sobj <- merged_obj_list_Sub[[obj]]
  # Step 1: Get the gene row sums across all barcodes (total count per gene)
  count_matrix <- sobj@assays$RNA$counts
  nCount_per_gene <- rowSums(count_matrix)
 if(identical(names(nCount_per_gene), rownames(sobj))) {#TRUE
 sobj@assays$RNA@meta.data$nCount_per_gene <- nCount_per_gene}
  merged_obj_list_Sub[[obj]] <- sobj
}

all_counts <- list()

for (obj in names(merged_obj_list_Sub)) {
  sobj <- merged_obj_list_Sub[[obj]]
  platform <- unique(sobj$Platform)
  Site <- unique(sobj$Site)
  Donor <- unique(sobj$Donor)
  Donor_Site <- unique(sobj$Donor_Site)
  gene_df <- sobj@assays$RNA@meta.data
  gene_df <- gene_df[!is.na(gene_df$percentage_gene_gc_content), ]

  gene_df <- gene_df %>%
  arrange(percentage_gene_gc_content)
  dens <- density(gene_df$percentage_gene_gc_content, na.rm = TRUE)
# 
# # Plot it to inspect
#  plot(dens, main = "Density Estimate of Gene GC percentage")
# abline(v = median(gene_df$percentage_gene_gc_content), col = "red", lty = 2)  # Add the raw data median

  # Create bins - use fewer bins for clarity, e.g. 100
  gene_df$GC_bin <- cut(
    gene_df$percentage_gene_gc_content,
    breaks = quantile(gene_df$percentage_gene_gc_content, probs = seq(0, 1, length.out = 100)),
    include.lowest = TRUE
  )

  counts_by_bin <- gene_df %>%
  group_by(GC_bin) %>%
  summarise(
    total_counts = sum(nCount_per_gene),
    gene_number = n()
  ) %>%
  mutate(
    platform = platform,
    Donor = Donor,
    Site = Site,
    Donor_Site = Donor_Site,
    Site_platform = paste0(Site, "_", platform),
    Sample_id = obj
  )
  all_counts[[obj]] <- counts_by_bin
}

# Combine all into one dataframe
combined_df <- bind_rows(all_counts)


scientific_num_pattern <- "[0-9\\.eE\\+\\-]+"
combined_df <- combined_df %>%
  mutate(
    bin_start = as.numeric(str_extract(GC_bin, paste0("(?<=\\[|\\()", scientific_num_pattern))),
    bin_end = as.numeric(str_extract(GC_bin, paste0("(?<=,)", scientific_num_pattern, "(?=\\]|\\))"))),
    bin_mid = (bin_start + bin_end) / 2
  )

combined_df$bin_mid_factor <- factor(combined_df$bin_mid)

combined_df <- combined_df %>%
  group_by(Sample_id) %>%
  mutate(
    total_counts_per_sample = sum(total_counts),  # Total counts for the entire sample
    # Normalize within each sample and calculate percentage of total counts per bin
    percentage_bin_count = (total_counts / total_counts_per_sample) * 100
  ) %>%
  ungroup()



pdf(file = paste0(fig_path_stnx, "07_Distribution_percentage_gene_gc_content_hvg_removed.pdf"), width = 8, height = 5)
ggplot(combined_df, aes(x = bin_mid_factor, y = percentage_bin_count, color = platform, group = Sample_id)) +
  geom_line(size = 0.5, alpha = 0.5) +
  labs(
    title = "Percentage of Counts per GC-Percentage-Bin per sample Across 11 Paired Samples \n MTs, RBs, HBs and Malat1 removed \n 50 bins ~720 genes per bin",
    x = "Gene GC-Percentage Bin Midpoint (bp)",
    y = "Percentage Counts per Bin per sample",
    color = "Sample"
  ) +
  scale_color_manual(values = c("scRNA" = "#00468B", "snRNA" = "#8B0046")) +
  scale_x_discrete(breaks = levels(combined_df$bin_mid_factor)[seq(1, length(levels(combined_df$bin_mid_factor)), by = 4)])+
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.line = element_line(color = "black")
  )


dev.off()
```
