---
title: "09_Celltype_Markers_Analysis"
author: "Reza Ghamsari"
date: "2025-06-10"
output: html_document
---
# Script Overview
This script performs comprehensive cell type marker analysis comparing single nuclei  vs single cell sequencing data.
It identifies marker genes for each cell type across samples, calculates Jaccard similarity indices to assess marker gene reproducibility,
and generates hierarchical clustering visualizations to evaluate method consistency.

```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now)
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

# Define Color Schemes and File Paths
```{r, colors}
# Custom color palette for consistent visualization
my_colors <- unique(c(
  "#802268FF", "#006400", "#525ecc99", "#1f0fdf99", "#480607", "#D55E00", 
  "#164194FF", "#7E6148FF", "#ADB6B6FF", "#52a7cc99", "#F0E685FF", "#cc52c099", 
  "#6acc5299", "#DC143C", "#802268FF", "#ffbb78", "#3C548899", "#189b3699", 
  "#cc9b5299", "#1f0fdf99", "#ccebc5", "#42857599", "#868686FF", "#006400", 
  "#0A47FFFF", "#3B1B53FF", "#749B58FF", "#ccc05299", "#cc765299", "#1B1919FF", 
  "#00D68FFF", "#14FFB1FF", "#3C5488FF", "#8F7700FF", "#164194FF", "#0094CDFF", 
  "#FFDAB9", "#FF00FF", "#00FFFF", "green", "blue", "#480607"
))

# Define the directory to save the PDF files
rds_path_vst <- "/vast/projects/aml_multiome/Sn_vs_SC/rds_path_vst/"
fig_path_stnx <- "/home/users/allstaff/ghamsari.r/seurat_figures/Sn_vs_SC_V4/"

```



# Loading Packages
```{r, pkgs}
script_title <- "09_Celltype_Markers_Analysis"
pkgs <- c("Signac", "Seurat", "BiocParallel", "DESeq2", "gridExtra", "parallel", "edgeR", "RColorBrewer", "patchwork", 
"limma", "tibble", "openxlsx", "ggplot2", "reshape2", "dplyr", "grid", "tidyverse", "htmlwidgets", "cowplot", "stringr", 
"tidyr", "pheatmap", "ggrepel", "future", "future.apply")


# Load all packages silently
invisible(lapply(pkgs, library, character.only = TRUE))

# Save session information for reproducibility
sink(file.path(fig_path_stnx, paste0(script_title, "_sessionInfo.txt")))
sessionInfo()
sink()
```

# Reading data
```{r, reading objects}
merged_obj <- readRDS(paste0(rds_path_vst, "05_Merge_object_cell_types_annotated_clustered_list_22_samples_11sn_11sc.rds"))

table(merged_obj$cell_type_broad,merged_obj$cell_type_fine,useNA = "ifany" )
table(merged_obj$cell_type_broad,merged_obj$sample_id,useNA = "ifany" )

```
# Marker Gene Identification
```{r, clustering}
# Idents(merged_obj) <- merged_obj$cell_type_broad
Idents(merged_obj) <- merged_obj$cell_type_fine
# Split object by sample for individual marker identification
merged_obj_list<- SplitObject(merged_obj, split.by = "sample_id")


# the number of available cores
num_cores <- parallelly::availableCores() #56
# Use multisession for stability (spawns separate R sessions instead of forking)
plan(multisession, workers = num_cores - 4)

FindAllMarkers_list <- function(x) {
  x <- NormalizeData(x, assay = "RNA")
  markers <- FindAllMarkers(x, 
                            only.pos = TRUE, 
                            min.cells.group = 3)
  return(markers)
}

all_markers_list <- future_lapply(merged_obj_list, 
                                  FindAllMarkers_list, 
                                  future.seed = TRUE)
# saveRDS(all_markers_list, paste0(rds_path_vst, "09_Find_all_markers_list_broad_cell_type.rds"))
saveRDS(all_markers_list, paste0(rds_path_vst, "09_Find_all_markers_list_cell_type_fine.rds"))
```

# Jaccard Similarity Analysis
Marker Gene Filtering and List Preparation

```{r, fig.width=20, fig.height=20}
# Extract top marker genes for each cell type in each sample
all_markers_list <- readRDS(paste0(rds_path_vst, "09_Find_all_markers_list_cell_type_fine.rds"))
# Initialize an empty list
filtered_gene_list <- list()

# Loop over each sample
for (sample_name in names(all_markers_list)) {
  marker_df <- all_markers_list[[sample_name]]
  
  # Optional: filter genes based on thresholds
  marker_df <- subset(marker_df, p_val_adj < 0.05 & avg_log2FC > 1)
  
  # Loop over each cluster (cell type)
  for (ct in unique(marker_df$cluster)) {
    df_ct <- marker_df[marker_df$cluster == ct, ]
    
    # Sort by adjusted p-value
    df_ct <- df_ct[order(df_ct$p_val_adj), ]
    
    genes <- df_ct$gene[1:200]
    print(paste0(sample_name, ";", ct , ": " , length(genes)))
    
    list_name <- paste0(sample_name, "__", ct)
    filtered_gene_list[[list_name]] <- unique(genes)
  }
}

#============================= Jaccard Index Calculation and Visualization
get_jaccard_matrix <- function(filtered_gene_list) {
  n <- length(filtered_gene_list)
  mat <- matrix(NA, n, n, dimnames = list(names(filtered_gene_list), names(filtered_gene_list)))
 for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      mat[i, j] <- length(intersect(filtered_gene_list[[i]], filtered_gene_list[[j]])) / 
                   length(union(filtered_gene_list[[i]], filtered_gene_list[[j]]))
    }
 }

  return(mat)
}

# Calculate Jaccard similarity matrix
mtx_jaccards <- get_jaccard_matrix(filtered_gene_list)
dim(mtx_jaccards) #330

# Convert upper triangle to full symmetric matrix
mtx_jaccards[lower.tri(mtx_jaccards)] <- t(mtx_jaccards)[lower.tri(mtx_jaccards)]


my_annotation_colors <- list(
  Method = c(cell = "#6699CC", Nuclei = "#CC6677"),
  Site = c(S1 = "#661100", S2 = "#888888", S3 = "#008B45", S4 = "#D55E00", ALL= "white" ),
  Donor = c(D1 = "#00A087FF", D2 = "#4A0F5FFF", D3 = "#00468B", D4 = "#BB0021FF",
            D5 = "#006400", D6 = "#08306B", D7 = "#ffbb78", D8 = "#8da0cb",
            D9 = "#e78ac3"),
  # cell_type = c("HSC" = "#ADB6B6FF",
  #                     "Erythroblast" = "#A20056FF",
  #                      "Reticulocyte" = "#BB0021FF",
  #                       "Normoblast" = "#9d183299",
  #                   "Monocytes" = "#D55E00",
  #                     "G.M_prog" = "#ffbb78",
  #                        "Lymph_Prog" = "#802268FF",
  #                   "pDC" = "#4A0F5FFF",
  #                     "B_Cells" = "#00468B",
  #                            "Plasmablast" = "#6699CC",
  #                     "NK_Cells" = "#008B45",
  #                     "T_cells" = "#00A087FF")
  cell_type =  c("HSC" = "#ADB6B6FF",
                      "Erythroblast" = "#A20056FF",
                      "Reticulocyte" = "#BB0021FF",
                       "Normoblast" = "#9d183299",
                        "CD16_Monocyte" = "#F05C3BFF",
                        "CD14_Monocyte" = "#D55E00",
                       "G.M_prog" = "#ffbb78",
                      "Monocyte_low_qc" = "#F0E685FF",
                         "Lymph_Prog" = "#802268FF",
                       "pDC" = "#4A0F5FFF",
                      "Transitianl_B_Cells" = "#08306B",
                      "B_Cells" = "#00468B",
                      "Plasmablast" = "#6699CC",
                      "NK_Cells" = "#008B45",
                      "CD4_T_cells" = "#00A087FF",
                      "CD8_T_cells" = "#91D1C2FF")
  )



length(rownames(mtx_jaccards))

# Create a metadata data.frame
meta <- data.frame(Sample = rownames(mtx_jaccards)) %>%
  separate(Sample, into = c("sample_id" , "cell_type"), sep = "__",  remove = FALSE) %>%
  separate(sample_id, into = c("Site", "Donor", "Method"), sep = "_",  remove = FALSE)


# Use as annotation row/column with reordered columns
annotation_col <- meta %>%
  select(Sample, cell_type, Method) %>%  # re-order columns
  column_to_rownames("Sample")

annotation_row <- meta %>%
  select(Sample,cell_type, Method, Site, Donor)%>%  # re-order columns
  column_to_rownames("Sample")

# matrix is a similarity matrix (Jaccard index between 0 and 1), not a distance matrix.Most clustering functions expect a distance matrix, so need to be converted:

dist_jaccard <- as.dist(1 - mtx_jaccards)

# Embed into Euclidean space
mds_coords <- cmdscale(dist_jaccard, k = 16) 

# Compute Euclidean distances on embedded data
dist_euc <- dist(mds_coords)

# Ward.D2 clustering on Euclidean distances
hc_ward <- hclust(dist_euc, method = "ward.D2")
# png_file <- paste0(fig_path_stnx, "09_Jaccard_markers_cell_type_fine_lfc_1_top200_large.png")
png_file <- paste0(fig_path_stnx, "09_Jaccard_markers_cell_type_fine_lfc_1_top200_small.png")
png(png_file, width = 9*600, height = 5*600, res = 600)
# png(png_file, width = 20*600, height = 20*600, res = 600)
# color_vector <- c("white", "#ad5fad")
color_vector <- c("white", "#737373")
fontsize = 5

# Use hc_ward in heatmap
pheatmap(mtx_jaccards,
         cluster_rows = hc_ward,
         cluster_cols = hc_ward,
         clustering_distance_rows = NULL,
         clustering_distance_cols = NULL,
         clustering_method = NULL,
         annotation_col = annotation_col,
         annotation_row = annotation_row,
         annotation_colors = my_annotation_colors,
         fontsize_row = fontsize,
         fontsize_col = fontsize,
         main = "Ward.D2 clustering on MDS-embedded Jaccard distances Jaccard Similarity of marker genes for each cell type within each sample across both methods",
         color = colorRampPalette(color_vector)(300),
         treeheight_col = 30,
         treeheight_row = 0)

dev.off()
# pheatmap(mtx_jaccards,
#          clustering_distance_rows = dist_jaccard,
# clustering_distance_cols = dist_jaccard,
#          # clustering_method = "complete",
# # Ward requires Euclidean distance because it minimizes within-cluster variance.
# # Jaccard distance violates this assumption — using Ward on it leads to poor or meaningless clusters.
#  clustering_method = "average",
#           cluster_rows = TRUE,
#          cluster_cols = TRUE,
#          annotation_col = annotation_col,
#          annotation_row = annotation_row,
#         annotation_colors = my_annotation_colors,
# fontsize_row = fontsize, fontsize_col = fontsize,
#          main = "Jaccard Similarity of marker genes for each cell type within each sample across both methods",
# color <- colorRampPalette(color_vector)(300),
#   # treeheight_row = 200,
# cutree_rows	= NA,
# cutree_cols = NA,
# 
#          treeheight_col = 100
# )


```

