---
title: "02_02_Batch_Pre_processing_multiome_QC_features"
author: "Reza Ghamsari"
date: "2024-08-26"
output: html_document
---

```{r setup, include=FALSE}
all_times <- list()  
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now)
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```
# Description

This script adds quality control metrics to each sample's Seurat object prior to filtering. Due to the size of the dataset (22 samples), processing was distributed across six SLURM runs, each handling a sub-list of samples. For each sample, RNA-based QC metrics are calculated including mitochondrial, ribosomal, MALAT1, haemoglobin, and lineage marker gene percentages, as well as chromosome-level expression proportions (chrX, chrY, chr7). ATAC-based metrics including nucleosome signal, TSS enrichment, ENCODE blacklist fraction, and per-chromosome peak percentages are also computed. The annotated sub-lists are saved separately and loaded together in the following script.


# Colors 
```{r colors}
my_colors <-c("#9d183299","#00A087FF", "#525ecc99","#E9967A", "#4DBBD5FF",  "#003C67FF", "#386cb0", "#7E6148FF",  "#ADB6B6FF", "#52a7cc99", "#F0E685FF", "#cc52c099",
              "#6acc5299","#DC143C", "#802268FF" ,"#ffbb78", "#3C548899", "#189b3699", "#cc9b5299", "#1f0fdf99", "#ccebc5", "#cc527699", "#42857599", "#868686FF"
              , "#006400", "#0A47FFFF" , "#3B1B53FF",  "#749B58FF" ,"#ccc05299", "#cc765299" , "#1B1919FF", "#00D68FFF" ,"#14FFB1FF","#3C5488FF" , "#8F7700FF" 
              , "#164194FF" , "#0094CDFF" , "#FFDAB9", "#FF00FF","#00FFFF","green" , "blue",  "#480607")

```
# Configuration
```{r config}
project_root <- here::here()
data_dir <- file.path(project_root, "data")
output_dir <- file.path(project_root, "outputs")
figures_dir <- file.path(project_root, "figures")

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(figures_dir, recursive = TRUE, showWarnings = FALSE)

```

# Loading Packages
```{r pkgs}
script_title <- "02_02_Batch_Pre_processing_multiome_QC_features"
packages <- c(
  "scales", "Seurat", "Signac", "ggplot2", "scCustomize", "gridExtra", 
  "parallelly", "parallel", "BiocParallel", "clustree", "RColorBrewer", 
  "patchwork", "igraph", "tibble", "ddqcR", "tidyr", "dplyr", "tidyverse", 
  "forcats", "Matrix", "EnsDb.Hsapiens.v86", "org.Hs.eg.db", "AnnotationDbi", 
  "biomaRt", "scDblFinder", "harmony", "pheatmap", "cowplot"
)
invisible(lapply(packages, library, character.only = TRUE))
sink(file.path(figures_dir,paste0(script_title ,"_sessionInfo.txt")))
sessionInfo()
sink()
```


# Reading data
```{r reading data}
multiome_list_11AML_9_healthy_2AMS <- readRDS(file.path(data_dir, "01_02_merged_multiome_11AML_9_healthy_2AMS_before_doublet_and_qc_not_merged_v1.rds"))
```

# split the merged multiome data
 The full list was too large to process in a single session, so it was split into 6 sub-lists,
 each processed in a separate SLURM run. QC-passed barcodes are later used to subset the merged object.
```{r split_batches}
length(multiome_list_11AML_9_healthy_2AMS) #22
# Seurat_merge_mutliome <- NULL
multiome_list_1_3 <- multiome_list_11AML_9_healthy_2AMS[1:3]
names(multiome_list_1_3)
multiome_list_4_7 <- multiome_list_11AML_9_healthy_2AMS[4:7]
names(multiome_list_4_7)
multiome_list_8_11 <- multiome_list_11AML_9_healthy_2AMS[8:11]
names(multiome_list_8_11)
multiome_list_12_15 <- multiome_list_11AML_9_healthy_2AMS[12:15]
names(multiome_list_12_15)
multiome_list_16_19 <- multiome_list_11AML_9_healthy_2AMS[16:19]
names(multiome_list_16_19)
multiome_list_20_22 <- multiome_list_11AML_9_healthy_2AMS[20:22]
names(multiome_list_20_22)

```


## Basic QC annotations
Each batch was processed in a separate SLURM run.
Uncomment the relevant batch_Number and multiome_list assignment before each run.
```{r QC_annotations, fig.width=20}
# #this is not a for loop, only change batch number in each run of slurm:
# batch_Number=1
# multiome_list <- multiome_list_1_3

# batch_Number=2
# multiome_list <- multiome_list_4_7

# batch_Number=4
# multiome_list <- multiome_list_8_11

# batch_Number=3
# multiome_list <- multiome_list_12_15

# batch_Number=5
# multiome_list <- multiome_list_16_19

batch_Number=6
multiome_list <- multiome_list_20_22
# Reading Seurat lists
batch_names <- c("list_1_3", "list_4_7", "list_8_11", "list_12_15", "list_16_19", "list_20_22")
batch_name <- batch_names[batch_Number]
print(batch_name)
# biomart
ensembl <- useMart("ensembl", dataset="hsapiens_gene_ensembl")
chr7_genes <- getBM(
  attributes = c('hgnc_symbol', 'chromosome_name'),
  filters = 'chromosome_name',
  values = '7',
  mart = ensembl
)
chr7_genes <- chr7_genes$hgnc_symbol
length(chr7_genes) # 1981
chr7_genes <- chr7_genes[chr7_genes%in%rownames(multiome_list[[1]])]
length(chr7_genes) # 1062
###########
chrX_genes <- getBM(
  attributes = c('hgnc_symbol', 'chromosome_name'),
  filters = 'chromosome_name',
  values = 'X',
  mart = ensembl
)
chrX_genes <- chrX_genes$hgnc_symbol
length(chrX_genes) # 1881
chrX_genes <- chrX_genes[chrX_genes%in%rownames(multiome_list[[1]])]
length(chrX_genes) # 926
#####
chrY_genes <- getBM(
  attributes = c('hgnc_symbol', 'chromosome_name'),
  filters = 'chromosome_name',
  values = 'Y',
  mart = ensembl
)
chrY_genes <- chrY_genes$hgnc_symbol
length(chrY_genes) # 469
chrY_genes <- chrY_genes[chrY_genes%in%rownames(multiome_list[[1]])]
length(chrY_genes) # 111
chrMT_genes <- getBM(
  attributes = c('hgnc_symbol', 'chromosome_name'),
  filters = 'chromosome_name',
  values = 'MT',
  mart = ensembl
)
chrMT_genes <- chrMT_genes$hgnc_symbol
length(chrMT_genes) # 37
chrMT_genes <- chrMT_genes[chrMT_genes%in%rownames(multiome_list[[1]])]
length(chrMT_genes) # 13

#The genes HBB,HBA2 are both haemoglobin genes and so should only be expressed in red blood cells and nowhere else. IGKC is an antibody gene produced only by B cells.
# MALAT1 is a long non-coding RNA that is predominantly localized in the nucleus, particularly in nuclear speckles which are subnuclear structures involved in the production and processing of pre-mRNA. Because of this localization, it would be expected that the signal for MALAT1 would be stronger in snRNA-seq data compared to scRNA-seq data, where the total RNA from the whole cell is being sequenced and the nuclear signal might be somewhat diluted by the presence of cytoplasmic RNAs
featureList <- list (
  # Quality Control Markers
  percentage_MT = "^MT-",                  # Mitochondrial genes
  percentage_RB = "^RP[SL]",               # Ribosomal genes
  percentage_Malat1 = "MALAT1",            # Nuclear-enriched non-coding RNA
  percentage_HB = "^HB[A-Z]*",
  # Hematopoietic Lineage Markers
  percentage_redblood.HBB_HBA2 = "^HB(B|A2)$",     # Red blood cell markers
  percentage_plat= "PECAM1|PF4",          # Platelet markers
  percentage_CD117 = "KIT",                # Hematopoietic stem cell marker
  percentage_CD33 = "CD33",                # Myeloid cell marker
  percentage_CD34 = "CD34",                # Hematopoietic stem/progenitor cell marker
  percentage_FLT3 = "FLT3",                # Hematopoietic stem/progenitor cell marker
  percentage_NPM1 = "NPM1",                # Leukemia marker
  percentage_t_cells = "CD3E|CD4|CD8A",            # T cell markers
  percentage_monocytes = "CD14|CD68",              # Monocyte/macrophage markers
  percentage_granulocytes = "CSF3R|FCGR3B",        # Granulocyte markers
  percentage_B_cell_markers = "MS4A1|CD79A|IGKC",       # B cell-specific markers
  percentage_NK_cell_markers = "NCAM1",            # NK cell marker
  percentage_endothelial_cells = "CDH5|VWF",       # Endothelial cell markers
  percentage_erythroid_cells = "GYPA|HBA1|HBB",    # Erythroid cell markers
  percentage_megakaryocytes = "ITGA2B|CD41",       # Megakaryocyte markers
  # AML Associated Markers
  percentage_CEBPA = "CEBPA",
  percentage_RUNX1 = "RUNX1",
  percentage_MLL = "KMT2A|MLL",
  percentage_TP53 = "TP53",
  percentage_IDH1 = "IDH1",
  percentage_IDH2 = "IDH2",
  percentage_TET2 = "TET2",
  percentage_CBX7 = "CBX7",
  percentage_DNMT3A = "DNMT3A",
  # Cell Survival Markers
  percentage_BCL2 = "BCL2",                        # Anti-apoptotic protein
  percentage_BCL2L1 = "BCL2L1",                    # Anti-apoptotic protein
  percentage_MCL1 = "MCL1",                        # Anti-apoptotic protein
  percentage_Ig_variable_markers = "^IGHV|^IGKV|^IGLV|^TRAV|^TRBV|^TRDV|^TRGV", #these are very patinet spacific so probably must remove from the data
  percentage_Ig_markers = "^IG[A-Z]*"
)

PercentageFeatures <- function(seuratObj) {
  DefaultAssay(seuratObj) <- "RNA"
for (featureName in names(featureList)) {
    patterns <- featureList[[featureName]]
    seuratObj[[featureName]] <- Seurat::PercentageFeatureSet(seuratObj, pattern = patterns)
}
#***********************************************************************
seuratObj[["percentage_chr7_genes"]] <- Seurat::PercentageFeatureSet(seuratObj, features = chr7_genes)
seuratObj[["percentage_chrMT_genes"]] <- Seurat::PercentageFeatureSet(seuratObj, features = chrMT_genes)
seuratObj[["percentage_chrY_genes"]] <- Seurat::PercentageFeatureSet(seuratObj, features = chrY_genes)
seuratObj[["percentage_chrX_genes"]] <- Seurat::PercentageFeatureSet(seuratObj, features = chrX_genes)

DefaultAssay(seuratObj) <- "ATAC"
seuratObj <- NucleosomeSignal(object = seuratObj)
seuratObj <- TSSEnrichment(object = seuratObj, fast = TRUE)
seuratObj$blacklist_fraction <- FractionCountsInRegion( object = seuratObj, assay = 'ATAC',regions = blacklist_hg38)
seuratObj$percentage_chr7_peaks_temp <- Seurat::PercentageFeatureSet(seuratObj, pattern = "chr7")

for (i in c(1:22, "X", "Y")) {
  chr_pattern <- paste0("chr", i)
  print(chr_pattern)
  feature_name <- paste0("percentage_chr", i, "_peaks")
  seuratObj[[feature_name]] <- Seurat::PercentageFeatureSet(seuratObj, pattern = chr_pattern)
}

  return(seuratObj)
}

multiome_list <- lapply(multiome_list, PercentageFeatures)
saveRDS(multiome_list, file.path(data_dir, paste0("02_02_multiome_feature_pre_", batch_name, "_v1.rds"))) #saved as different batches
```




