---
title: "03_Batch_Pre_processing_multiome_QC_features""
author: "Reza Ghamsari"
date: "2024-09"
output: html_document
---
```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now)
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```
#Description
This script performs doublet detection and multi-step quality control filtering across all 22 samples (11 AML, 9 healthy BMMC, 2 AMS). After removing confirmed doublets, cells are filtered using fixed QC thresholds followed by a cluster-aware MAD-based outlier detection step. C Sample-specific features including chrX, chrY, chr7 genes and peaks, immunoglobulin/TCR variable genes, and haemoglobin genes are also removed from the final merged object.

# Colors 
```{r colors}
my_colors <-c("#9d183299","#00A087FF", "#525ecc99","#E9967A", "#4DBBD5FF",  "#003C67FF", "#386cb0", "#7E6148FF",  "#ADB6B6FF"
              , "#52a7cc99", "#F0E685FF", "#cc52c099",  "#6acc5299","#DC143C", "#802268FF" ,"#ffbb78", "#3C548899", "#189b3699",
              "#cc9b5299", "#1f0fdf99", "#ccebc5", "#cc527699", "#42857599", "#868686FF", "#006400", "#0A47FFFF" , "#3B1B53FF", 
              "#749B58FF" ,"#ccc05299", "#cc765299" , "#1B1919FF", "#00D68FFF" ,"#14FFB1FF","#3C5488FF" , "#8F7700FF" , "#164194FF" ,
              "#0094CDFF" , "#FFDAB9", "#FF00FF","#00FFFF","green" , "blue",  "#480607")
```
# Configuration
```{r config}
project_root <- here::here()
data_dir <- file.path(project_root, "data")
output_dir <- file.path(project_root, "outputs")
figures_dir <- file.path(project_root, "figures")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(figures_dir, recursive = TRUE, showWarnings = FALSE)
```


# Loading Packages
```{r pkgs}
script_title <- "03_Batch_Pre_processing_multiome_QC_features"
packages <- c(
  "scales", "Seurat", "Signac", "ggplot2", "scCustomize", "gridExtra", 
  "parallelly", "parallel", "BiocParallel", "clustree", "RColorBrewer", 
  "patchwork", "igraph", "tibble", "ddqcR", "tidyr", "dplyr", "tidyverse", 
  "forcats", "Matrix", "EnsDb.Hsapiens.v86", "org.Hs.eg.db", "AnnotationDbi", 
  "biomaRt", "scDblFinder", "harmony", "pheatmap", "cowplot"
)

invisible(lapply(packages, library, character.only = TRUE))
sink(file.path(figures_dir, paste0(script_title ,"_sessionInfo.txt")))
sessionInfo()
sink()
```

After adding different qc features for each object across different batches in the previous script, I will now use scdbFinder to detect doublets. 
I'm performing clustering before conducting quality control and removing outliers, based on the assumption that doublets behave like outliers. 
By retaining them during clustering, I aim to improve scdbFinder's ability to identify these outliers more effectively.


```{r reading object}
multiome_list <- list()
batch_names <- c("list_1_3", "list_4_7", "list_8_11", "list_12_15", "list_16_19", "list_20_22")

for (batch_name in batch_names){
  print(batch_name)
 batch_list <- readRDS(file.path(output_dir,paste0("02_02_multiome_feature_pre_", batch_name, "_v1.rds")))
 multiome_list <- c(multiome_list, batch_list)
}
length(multiome_list)
```

# Second round of Preprocessing for Doublet detection
```{r clustering, fig.width=10}

identical(multiome_list$Sample_1@assays$RNA@data, multiome_list$Sample_1@assays$RNA@counts) #TRUE
identical(multiome_list$Sample_1@assays$ATAC@data, multiome_list$Sample_1@assays$ATAC@counts) #TRUE

NormReduc <- function(x) {
  DefaultAssay(x) <- "RNA"
  x <- SCTransform(x, verbose = FALSE, vst.flavor = "v2", variable.features.n = 3000,new.assay.name= "SCT_doub") 
  x <- RunPCA(x, reduction.name='pca_sct_doub', assay= "SCT_doub")
  DefaultAssay(x) <- "ATAC"
  x <- RunTFIDF(x)
  x <- FindTopFeatures(x, min.cutoff = 'q25')
  x <- RunSVD(x, reduction.name= 'lsi_doub',  reduction.key = "doub_LSI_")
  x<- FindMultiModalNeighbors(x, reduction.list = list("pca_sct_doub", "lsi_doub"), dims.list = list(c(1:50), c(1:50)), snn.graph.name="wsnn_doub",  knn.graph.name = "wknn_doub",  weighted.nn.name = "weighted_doub.nn")
  }

multiome_list <- mclapply(multiome_list, NormReduc, mc.cores = detectCores() - 5, mc.preschedule = FALSE)
```
# Clustering
```{r Clustering}

resolutions <- unique(c(0.001, 0.0025, 0.015, 0.045, seq(0.005, 0.1, 0.01), seq(0.1, 1.5, 0.2), seq(2, 3, 0.5)))

 ClusterParallel <- function(x, graph_name) {
  graph <- x[[graph_name]]
  x <- Seurat:::FindClusters.default(
  object = graph, 
  algorithm = 1, 
  verbose = TRUE, 
  random.seed = 0, 
  resolution = resolutions, 
  n.iter = 10
)
  return(x)
}

clustering_list <- mclapply(multiome_list, ClusterParallel, graph_name = "wsnn_doub", mc.cores = detectCores() - 5, mc.preschedule = FALSE)


```

## Clustree
```{r clustree, fig.width=20}
ClustreeParallel <- function(x, name) {
  plot <- clustree(x, prefix = "res.", node_label_size = 4, node_text_size = 5, 
                   node_alpha = 0.3, scale_node_text = FALSE, 
                   node_size_range = c(4, 20)) +
    scale_color_manual(values = my_colors) +
    guides(color = guide_legend(ncol = 2)) +
    ggtitle(paste("Clustree", name))
  
  return(plot)
}

page_plots <- mclapply(seq_along(clustering_list), function(i) {
  ClustreeParallel(clustering_list[[i]], names(clustering_list)[i])
}, mc.cores = detectCores() - 5, mc.preschedule = FALSE)

names(page_plots) <- names(clustering_list)



pdf(file = file.path(figures_dir, "03_Clustree_wnn_11AML_9healthy_2AMS_samples_before_doublet_detection_before_QC.pdf"), width = 30, height = 25,  pointsize = 5) 
for (plot in page_plots) {
  print(plot)
}
dev.off()
```

## cluster annotation:
based on best stability in clustree visualization, I chose the best resolution for each sample.
```{r}
# Manually assigning clustering values to multiome_list for each sample
multiome_list$Sample_1$cluster_doublet <- clustering_list$Sample_1$res.1.1
multiome_list$Sample_3$cluster_doublet <- clustering_list$Sample_3$res.1.1
multiome_list$Sample_5$cluster_doublet <- clustering_list$Sample_5$res.0.9
multiome_list$Sample_7$cluster_doublet <- clustering_list$Sample_7$res.0.9
multiome_list$Sample_10$cluster_doublet <- clustering_list$Sample_10$res.0.9
multiome_list$Sample_13$cluster_doublet <- clustering_list$Sample_13$res.1.1
multiome_list$Sample_16$cluster_doublet <- clustering_list$Sample_16$res.0.9
multiome_list$Sample_18$cluster_doublet <- clustering_list$Sample_18$res.1.1
multiome_list$Sample_20$cluster_doublet <- clustering_list$Sample_20$res.1.5
multiome_list$Sample_22$cluster_doublet <- clustering_list$Sample_22$res.1.1
multiome_list$Sample_24$cluster_doublet <- clustering_list$Sample_24$res.1.1
multiome_list$BMMC_S1_D1$cluster_doublet <- clustering_list$BMMC_S1_D1$res.1.1
multiome_list$BMMC_S1_D2$cluster_doublet <- clustering_list$BMMC_S1_D2$res.0.7
multiome_list$BMMC_S1_D3$cluster_doublet <- clustering_list$BMMC_S1_D3$res.1.3
multiome_list$BMMC_S2_D1$cluster_doublet <- clustering_list$BMMC_S2_D1$res.1.3
multiome_list$BMMC_S3_D6$cluster_doublet <- clustering_list$BMMC_S3_D6$res.1.5
multiome_list$BMMC_S3_D7$cluster_doublet <- clustering_list$BMMC_S3_D7$res.0.9
multiome_list$BMMC_S4_D1$cluster_doublet <- clustering_list$BMMC_S4_D1$res.0.5
multiome_list$BMMC_S4_D8$cluster_doublet <- clustering_list$BMMC_S4_D8$res.0.5
multiome_list$BMMC_S4_D9$cluster_doublet <- clustering_list$BMMC_S4_D9$res.0.9
multiome_list$`6279DN`$cluster_doublet <- clustering_list$`6279DN`$res.0.9
multiome_list$`6905DN`$cluster_doublet <- clustering_list$`6905DN`$res.1.3
```

# multiplet rate
Here I compare scDblFinder's default `dbr` calculation method with the 10x Multiome manual multiplet rate. 
For samples with high cell counts, the default method appeared to overestimate doublet rates, 
so I used a modified exponential model fitted to the 10x protocol data instead.

```{r}

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@  sdblfinder default dbr calculation method
calculate_dbr <- function(cell_counts) {
  dbr <- weighted.mean(cell_counts / 100000, cell_counts)
  return(dbr)
}

num_cells <- seq(1000, 20000, 1000)

dbr_values <- sapply(num_cells, calculate_dbr)


plot_data <- data.frame(CellCounts = num_cells, DBR = dbr_values)


ggplot(plot_data, aes(x = CellCounts, y = DBR)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "red", size = 2) +
  labs(title = "DBR vs. Cell Counts/ based on weighted.mean(cell_counts / 100000, cell_counts) \n scDblFinder",
       x = "Cell Counts",
       y = "DBR") +
  theme_minimal()

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@  10x protocol data

# Defining the data from the table
multiplet_rate <- c(0.004, 0.008, 0.016, 0.023, 0.031, 0.039, 0.046, 0.054, 0.062, 0.069, 0.077, 0.12, 0.16)
targeted_nuclei <- c(500, 1000, 2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000, 15000, 20000)

# Fit an exponential model
exponential_model <- nls(multiplet_rate ~ a * exp(b * targeted_nuclei), start = list(a = 0.004, b = 0.0005))
model_params <- summary(exponential_model)$coefficients

# Defining the function to calculate multiplet rate based on the fitted model
calculate_multiplet_rate <- function(targeted_nuclei) {
  a <- model_params[1, 1]
  b <- model_params[2, 1]
  multiplet_rate <- a * exp(b * targeted_nuclei)
  return(multiplet_rate)
}

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# Generate nuclei counts for plotting
nuclei_counts <- seq(500, 20000, 500)
fitted_multiplet_rates <- calculate_multiplet_rate(nuclei_counts)
plot_data <- data.frame(TargetedNuclei = targeted_nuclei, MultipletRate = multiplet_rate)
fitted_data <- data.frame(TargetedNuclei = nuclei_counts, MultipletRate = fitted_multiplet_rates)
ggplot() +
  geom_point(data = plot_data, aes(x = TargetedNuclei, y = MultipletRate), color = "red", size = 2) +
  geom_line(data = fitted_data, aes(x = TargetedNuclei, y = MultipletRate), color = "blue", size = 1) +
  labs(title = "Modified dbr based on 10x protocol \n multiplet_rate <- a * exp(b * targeted_nuclei)",
       x = "Nuclei Counts",
       y = "DBR") +
  theme_minimal()



#@@@@@@@@@@@@@@@@@@@@@@@@@@@@  comparison of my modified dbr and default one 
results <- data.frame(
  SampleName = character(),
  CellNumber = numeric(),
  DBR_Modified = numeric(),
  DBR_ScDblFinder = numeric(),
  stringsAsFactors = FALSE
)

# Looping through the samples in multiome_list and calculate the doublet rates
for (sobj in names(multiome_list)) {
  sobject <- multiome_list[[sobj]]
  sample_Name <- sobj
  cell_number <- ncol(sobject)
  dbr_modified <- calculate_multiplet_rate(cell_number)
  dbr_scdbfinder <- calculate_dbr(cell_number)
 
   results <- rbind(results, data.frame(
    SampleName = sample_Name,
    CellNumber = cell_number,
    DBR_Modified = dbr_modified,
    DBR_ScDblFinder = dbr_scdbfinder
  ))
}

results <- results[order(results$DBR_Modified), ]

# Converting SampleName to a factor to retain the order in the plot
results$SampleName <- factor(results$SampleName, levels = results$SampleName)

print(results)

ggplot(results, aes(x = SampleName)) +
  geom_line(aes(y = DBR_Modified, color = "Modified DBR", group = 1)) +
  geom_line(aes(y = DBR_ScDblFinder, color = "ScDblFinder DBR", group = 1)) +
  geom_point(aes(y = DBR_Modified, color = "Modified DBR")) +
  geom_point(aes(y = DBR_ScDblFinder, color = "ScDblFinder DBR")) +
  labs(title = "Doublet Rate Comparison",
       x = "Sample Name",
       y = "Doublet Rate",
       color = "Legend") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))  

```


# RNA Doublet detection
I used modified dbr here

```{r, RNA Doublet detection}
bp <- MulticoreParam(3, RNGseed=1234) #Because it relies on the partly random generation of artificial doublets, running scDblFinder multiple times on the same data will yield slightly different results. 
# We can ensure reproducibility using set.seed(123), however this will not be sufficient when processing multiple samples (i.e. using the samples argument – even without multithreading!). In such case, the seed needs to be passed to the BPPARAMs:
# BPPARAM=bp)

# Function for doublet finder 
DOUBLET_detect <- function(sample) {
  db <- calculate_multiplet_rate(ncol(sample))
  scDblFinder::scDblFinder(Seurat::GetAssayData(sample, slot = "counts", assay = "RNA"), 
                   clusters = sample$cluster_doublet,
                   processing="default", 
                     returnType = c("sce", "table", "full", "counts"),
                   verbose = TRUE,
                  dbr= db,
                   BPPARAM = bp)
}

RNA_doublet_list_scdbfinder <- purrr::map(multiome_list, DOUBLET_detect)
MAKE_TABLE <- function(obj, id) {
  df <- as.data.frame(table(obj$scDblFinder.class))
  df$Sample <- id
  return(df)
}


#Iterateing over names of the list for Sample IDs
combined_df <- purrr::map2_df(RNA_doublet_list_scdbfinder, names(RNA_doublet_list_scdbfinder), MAKE_TABLE)
reshaped_df <- combined_df %>%
  spread(key = Var1, value = Freq)
reshaped_df <- reshaped_df %>%
  mutate(doublet_percentage = 100 * doublet / (doublet + singlet))


write.csv(reshaped_df, file = file.path(figures_dir, "03_Doublet_percentage__RNA_11AML_9healthy_2AMS_modified_md_dbr.csv"), row.names = T)
reshaped_df$Total_cell <- reshaped_df$singlet +reshaped_df$doublet

 

file_path <- file.path(figures_dir, "03_Doublet_percentage__RNA_11AML_9healthy_2AMS_modified_md_dbr.png")

png(file_path, width = 2000, height = 1000, res = 130)

# Setting up the layout for two plots side by side
par(mfrow = c(1, 2))

plot(reshaped_df$Total_cell, reshaped_df$doublet_percentage,
     xlab = "Total Cell", ylab = "Doublet Percentage",
     main = "Doublet Percentage vs Total number of Nuclei (22 samples, unfiltered, md_dbr)")
abline(lm(doublet_percentage ~ Total_cell, data = reshaped_df), col = "red")

plot(reshaped_df$Total_cell, reshaped_df$doublet,
     xlab = "Total Cell", ylab = "Doublet number",
     main = "Doublet number vs Total number of Nuclei (22 samples, unfiltered, md_dbr)")
abline(lm(doublet ~ Total_cell, data = reshaped_df), col = "red")

dev.off()





for (i in names(multiome_list)) {
  for( j in names(RNA_doublet_list_scdbfinder)) {
    if (i == j) {
      print(paste("Checking identity for", i))
      is_identical <- identical(rownames(multiome_list[[i]]@meta.data),
                                colnames(RNA_doublet_list_scdbfinder[[j]]))
      print(is_identical)
    }
  }
}



for (i in names(multiome_list)) {
  for( j in names(RNA_doublet_list_scdbfinder)) {
    if (i == j) {
      multiome_list[[i]]@meta.data$scDblFinder_RNA_counts = RNA_doublet_list_scdbfinder[[j]]$scDblFinder.class
    }
  }
}


for (i in names(multiome_list)) {
  for( j in names(RNA_doublet_list_scdbfinder)) {
    if (i == j) {
      multiome_list[[i]]@meta.data$scDblFinder_RNA_counts_score = RNA_doublet_list_scdbfinder[[j]]$scDblFinder.score
    }
  }
}



for (i in names(multiome_list)) {
  for( j in names(RNA_doublet_list_scdbfinder)) {
    if (i == j) {
      print(paste("Checking identity for", i))
      is_identical <- identical(multiome_list[[i]]@meta.data$scDblFinder_RNA_counts,
                                RNA_doublet_list_scdbfinder[[j]]$scDblFinder.class)
      print(is_identical)
    }
  }
}


```



# ATAC Doublet detection
https://bioconductor.org/packages/release/bioc/vignettes/scDblFinder/inst/doc/scATAC.html

```{r, ATAC Doublet detection}
bp <- MulticoreParam(3, RNGseed=1234) 
DOUBLET_ATAC <- function(sample) {
  db <- calculate_multiplet_rate(ncol(sample))
  scDblFinder::scDblFinder(Seurat::GetAssayData(sample, slot = "counts", assay = "ATAC"), 
                   clusters = sample$cluster_doublet,
                  artificialDoublets=NULL,
                   aggregateFeatures=TRUE, 
                   nfeatures=25,
                  processing="rawFeatures",
                    returnType = c("sce", "table", "full", "counts"),
                  dbr= db,
                   verbose = TRUE,
                   BPPARAM = bp)
}


atac_doublet_list_scdbfinder <- purrr::map(multiome_list, DOUBLET_ATAC)

MAKE_TABLE <- function(obj, id) {
  df <- as.data.frame(table(obj$scDblFinder.class))
  df$Sample <- id
  return(df)
}


# Using map2 to also iterate over names of the list for Sample IDs
combined_df <- purrr::map2_df(atac_doublet_list_scdbfinder, names(atac_doublet_list_scdbfinder), MAKE_TABLE)


reshaped_df <- combined_df %>%
  spread(key = Var1, value = Freq)
reshaped_df <- reshaped_df %>%
  mutate(doublet_percentage = 100 * doublet / (doublet + singlet))
reshaped_df$Total_cells <- reshaped_df$singlet +reshaped_df$doublet

write.csv(reshaped_df, file = file.path(figures_dir, "03_Doublet_percentage_ATAC_11AML_9healthy_2AMS_modified_md_dbr.csv"), row.names = T)

```

#Integrating doublet annotations from ATAC and RNA
In this section, ATAC-based doublet detection results from **scDblFinder** are integrated 
 combined doublet classification (`scDblFinder_Doublets`) is created to distinguish:  
  - `Doublet_both` (detected in both RNA and ATAC),  
  - `Doublet_ATAC` (detected only in ATAC),  
  - `Doublet_RNA` (detected only in RNA),  
  - `singlet`.  
```{r integrating the annotation}
for (i in names(multiome_list)) {
  for( j in names(atac_doublet_list_scdbfinder)) {
    if (i == j) {
      print(paste("Checking identity for", i))
      is_identical <- identical(rownames(multiome_list[[i]]@meta.data),
                                colnames(atac_doublet_list_scdbfinder[[j]]))
      print(is_identical)
    }
  }
}



for (i in names(multiome_list)) {
  for( j in names(atac_doublet_list_scdbfinder)) {
    if (i == j) {
      multiome_list[[i]]@meta.data$scDblFinder_ATAC_counts = atac_doublet_list_scdbfinder[[j]]$scDblFinder.class
    }
  }
}

for (i in names(multiome_list)) {
  for( j in names(atac_doublet_list_scdbfinder)) {
    if (i == j) {
      multiome_list[[i]]@meta.data$scDblFinder_ATAC_counts_score = atac_doublet_list_scdbfinder[[j]]$scDblFinder.score
    }
  }
}

for (i in names(multiome_list)) {
  for( j in names(atac_doublet_list_scdbfinder)) {
    if (i == j) {
      print(paste("Checking identity for", i))
      is_identical <- identical(multiome_list[[i]]@meta.data$scDblFinder_ATAC_counts,
                                atac_doublet_list_scdbfinder[[j]]$scDblFinder.class)
      print(is_identical)
    }
  }
}


multiome_list <- lapply(multiome_list, function(x) {
  x@meta.data[["scDblFinder_Doublets"]] <- 
 ifelse (
   x@meta.data[["scDblFinder_ATAC_counts"]] == "doublet" & x@meta.data[["scDblFinder_RNA_counts"]] == "doublet", "Doublet_both" ,
   ifelse(
     x@meta.data[["scDblFinder_ATAC_counts"]] == "doublet" & x@meta.data[["scDblFinder_RNA_counts"]] == "singlet","Doublet_ATAC" ,
     ifelse(
       x@meta.data[["scDblFinder_RNA_counts"]] == "doublet" & x@meta.data[["scDblFinder_ATAC_counts"]] == "singlet","Doublet_RNA", 
       "singlet"
     )
   )
 )
  x
  })

table(multiome_list$Sample_10$scDblFinder_Doublets)

meta_data_list <- list()

for (donor in names(multiome_list)) {
  meta_data <- multiome_list[[donor]]@meta.data
  meta_data_list[[donor]] <- meta_data}

  
saveRDS(meta_data_list, file = file.path(output_dir, "03_meta_data_list_11AML_9healthy_2AMS_doublet_Removed_before_qc.rds"))

```


# removing cells that are doublet for both
```{r}
meta_data_list <- readRDS(file.path(output_dir, "03_meta_data_list_11AML_9healthy_2AMS_doublet_Removed_before_qc.rds"))
for (donor in names(multiome_list)) {
  meta_data <- meta_data_list[[donor]]
  multiome_list[[donor]]@meta.data <- meta_data}



for (sobj_name in names(multiome_list)) {
  sobj <- multiome_list[[sobj_name]]
  print(sobj_name)
  # sobj_name@assays$SCT_doub <- NULL #remove the extra assay to save space
  print(paste0("before filtering: ",length(colnames(sobj))))
  DefaultAssay(sobj) <- "RNA"
    sobj <- subset(
      x = sobj,
      subset = scDblFinder_Doublets != "Doublet_both"
    ) 
    
    # Update the Seurat object in the list
    print(paste0("After filtering: ", length(colnames(sobj))))
    multiome_list[[sobj_name]] <- sobj}

```
#removing outliers based on quantiles
To remove low-quality cells, we applied a soft QC threshold and Cells falling below the 0.2 percent quantile or above the 99.8 percent quantile for any feature are flagged as `"outlier"`. 
```{r  removing quantile}
lower_quantile <- 0.002
upper_quantile <- 0.998
features <- c("nCount_RNA", "nFeature_RNA", "nCount_ATAC", "nFeature_ATAC")

# A function to identify outliers for a sample
identify_sample_outliers <- function(sobj, features, lower_quantile, upper_quantile) {
  is_outlier <- rep("passed", nrow(sobj@meta.data))  # Initialised as "passed"
  
  for (feature in features) {
    feature_data <- sobj@meta.data[[feature]]
    lower_bound <- quantile(feature_data, probs = lower_quantile, na.rm = TRUE)
    upper_bound <- quantile(feature_data, probs = upper_quantile, na.rm = TRUE)
    is_outlier <- ifelse(is_outlier == "passed" & (feature_data < lower_bound | feature_data > upper_bound), "outlier", is_outlier)
  }
  return(is_outlier)
}

# Loop over samples
for (sample in names(multiome_list)) {
  print(sample)
  sobj <- multiome_list[[sample]]
  outliers <- identify_sample_outliers(sobj, features, lower_quantile, upper_quantile)
  print(table(outliers))
  multiome_list[[sample]]$outlier <- NULL
   multiome_list[[sample]]$quantile_004_percent <- outliers
}


table(multiome_list$Sample_10$quantile_004_percent)
# outlier  passed 
#      99   10409
```
# violin plot QC
```{r}

features <- c("nCount_RNA", "nFeature_RNA", "nCount_ATAC", "nFeature_ATAC", "percentage_MT", "percentage_HB" , "percentage_RB", "percentage_Malat1", "nucleosome_signal", "TSS.enrichment" , "blacklist_fraction", "nucleosome_signal",
              "percentage_chrY_genes", "percentage_chrX_genes",
  "percentage_chr7_peaks_temp", "percentage_chr1_peaks", "percentage_chr2_peaks", "percentage_chr3_peaks",
  "percentage_chr4_peaks", "percentage_chr5_peaks", "percentage_chr6_peaks", "percentage_chr7_peaks",
  "percentage_chr8_peaks", "percentage_chr9_peaks", "percentage_chr10_peaks", "percentage_chr11_peaks",
  "percentage_chr12_peaks", "percentage_chr13_peaks", "percentage_chr14_peaks", "percentage_chr15_peaks",
  "percentage_chr16_peaks", "percentage_chr17_peaks", "percentage_chr18_peaks", "percentage_chr19_peaks",
  "percentage_chr20_peaks", "percentage_chr21_peaks", "percentage_chr22_peaks", "percentage_chrX_peaks",
  "percentage_chrY_peaks", "percentage_chr7_genes", "percentage_chrMT_genes",

  "percentage_redblood.HBB_HBA2", "percentage_plat",
  "percentage_CD117", "percentage_CD33", "percentage_CD34", "percentage_FLT3", "percentage_NPM1", "percentage_t_cells",
  "percentage_monocytes", "percentage_granulocytes", "percentage_B_cell_markers", "percentage_NK_cell_markers",
  "percentage_endothelial_cells", "percentage_erythroid_cells", "percentage_megakaryocytes", "percentage_CEBPA",
  "percentage_RUNX1", "percentage_MLL", "percentage_TP53", "percentage_IDH1", "percentage_IDH2", "percentage_TET2",
  "percentage_CBX7", "percentage_DNMT3A", "percentage_BCL2", "percentage_BCL2L1", "percentage_MCL1", "percentage_Ig_variable_markers",
  "percentage_Ig_markers"
)

samples <- names(multiome_list)

pdf(file = file.path(figures_dir, "03_Violine_plot_multiome_list_11AML_9healthy_2AMS_doublet_Removed_before_qc_not_log.pdf") , width = 20, height = 10)
# pdf(file = file.path(figures_dir, "03_Violine_plot_multiome_list_11AML_9healthy_2AMS_doublet_Removed_before_qc_log.pdf") , width = 20, height = 10)

# Loop over each feature
for (feature in features) {
  combined_data <- do.call(rbind, lapply(samples, function(sample) {
    if (feature %in% colnames(multiome_list[[sample]]@meta.data)) {
      data.frame(Sample = sample, FeatureValue = multiome_list[[sample]]@meta.data[[feature]])
    }
  }))
  
  p <- ggplot(combined_data, aes(x = Sample, y = FeatureValue, fill = Sample)) + 
        geom_violin(trim = FALSE, color = NA)  +  
        geom_jitter(width = 0.4, size = 0.1, alpha = 0.1, color = "gray") +
      scale_fill_manual(values = my_colors) +
       # scale_y_log10() +  # 
       labs(title = paste("Violin Plot for", feature), 
            x = "Samples", 
            y = feature) +
       theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1),  
             panel.grid.major = element_blank(),  
             panel.grid.minor = element_blank())

  print(p)
}

dev.off()
```


# Subsetting filtered cells

To remove low-quality cells, we applied a soft QC threshold and Cells falling below the 0.2 percent quantile or above the 99.8 percent quantile for any feature are flagged as `"outlier"`. For ATAC, cells with 300–150,000 counts and 300–50,000 detected features were retained, while for RNA, cells with 200–100,000 counts and 200–6,000 detected genes were kept. Additional filters excluded cells with >10\% of reads mapping to MALAT1, >40% to ribosomal genes, >5% to mitochondrial genes, or >3% of fragments overlapping ENCODE blacklist regions. Cells with nucleosome signal more than 4 or transcription start site (TSS) enrichment less than 3 were also removed. 
```{r subsetting}
removed_cells_counts <- list()
for (sample in names(multiome_list)) {
  print(sample)
  sobj <- multiome_list[[sample]]
  col1 <- ncol(sobj)
  sobj <- subset(
  x = sobj,
  subset = 
    nCount_ATAC > 300 &
    nCount_ATAC < 150000 &
    nFeature_ATAC > 300 &
    nFeature_ATAC < 50000 &
    nCount_RNA > 200 &
    nCount_RNA < 100000 &
     nFeature_RNA < 6000 &
    nFeature_RNA > 200 &
    percentage_Malat1 < 10 &
    percentage_RB < 40 &
    percentage_MT < 5 &
    blacklist_fraction < 0.03 &
    nucleosome_signal < 4 & 
    TSS.enrichment > 3 &
    quantile_004_percent == "passed"

)

# visualisation
 multiome_list[[sample]] <- sobj
    col2 <- ncol(sobj)
    removed_cells_counts[[sample]] <- print(paste0("removed cells in ", sample," = ", col1-col2, "out of = " , col1, " cells."))
    print(paste0("removed cells in ", sample," = ", col1-col2))
}

# Save counts to a text file
write.table(removed_cells_counts, file.path(figures_dir, "03_first_removed_cells_based_on_violin_before_ddqc_v1.txt"), row.names = TRUE, col.names = TRUE)

```
# Biolody aware QC
```{r clustering, }
identical(multiome_list$Sample_1@assays$RNA@data, multiome_list$Sample_1@assays$RNA@counts) #TRUE
identical(multiome_list$Sample_1@assays$ATAC@data, multiome_list$Sample_1@assays$ATAC@counts) #TRUE
NormReduc <- function(x) {
  DefaultAssay(x) <- "RNA"
  x <- SCTransform(x, verbose = FALSE, vst.flavor = "v2", variable.features.n = 3000,new.assay.name= "SCT_QC") 
  x <- RunPCA(x, reduction.name='pca_sct_QC', assay= "SCT_QC")
  DefaultAssay(x) <- "ATAC"
x <- RunTFIDF(x)
x <- FindTopFeatures(x, min.cutoff = 'q25')
x <- RunSVD(x, reduction.name= 'lsi_QC',  reduction.key = "QC_LSI_")
x<- FindMultiModalNeighbors(x, reduction.list = list("pca_sct_QC", "lsi_QC"), dims.list = list(c(1:50), c(2:40)), snn.graph.name="wsnn_QC",  knn.graph.name = "wknn_QC",  weighted.nn.name = "weighted_QC.nn")

}

multiome_list <- mclapply(multiome_list, NormReduc, mc.cores = detectCores() - 10, mc.preschedule = FALSE)
for (sobj_name in names(multiome_list)) {
  # Retrieve the Seurat object
  sobj <- multiome_list[[sobj_name]]
  print(sobj_name)
  sobj@assays$SCT_QC <- NULL #remove the extra assay to save space
  multiome_list[[sobj_name]] <- sobj
    }
    
```


## clustering 
```{r, clustering}

resolutions <- unique(c(0.001, 0.0025, 0.015, 0.045, seq(0.005, 0.1, 0.01), seq(0.1, 1.5, 0.2), seq(2, 3, 0.5)))
 ClusterParallel <- function(x, graph_name) {
  graph <- x[[graph_name]]
  x <- Seurat:::FindClusters.default(
  object = graph, 
  algorithm = 1, 
  verbose = TRUE, 
  random.seed = 0, 
  resolution = resolutions, 
  n.iter = 10
)
  return(x)
}

clustering_list_QC <- mclapply(multiome_list, ClusterParallel, graph_name = "wsnn_QC", mc.cores = detectCores() - 5, mc.preschedule = FALSE)


```


## clustree
```{r}
ClustreeParallel <- function(x, name) {
  plot <- clustree(x, prefix = "res.", node_label_size = 4, node_text_size = 5, 
                   node_alpha = 0.3, scale_node_text = FALSE, 
                   node_size_range = c(4, 20)) +
    scale_color_manual(values = my_colors) +
    guides(color = guide_legend(ncol = 2)) +
    ggtitle(paste("Clustree", name))
  
  return(plot)
}

page_plots <- mclapply(seq_along(clustering_list_QC), function(i) {
  ClustreeParallel(clustering_list_QC[[i]], names(clustering_list_QC)[i])
}, mc.cores = detectCores() - 5, mc.preschedule = FALSE)

names(page_plots) <- names(clustering_list_QC)

 pdf(file = file.path(figures_dir, "03_Clustree_wnn_11AML_9healthy_2AMS_samples_after_doublet_and_preqc_removal_before_ddQC.pdf"), width = 30, height = 25,  pointsize = 5) 

for (plot in page_plots) {
  print(plot)
}
 
dev.off()

```
## cluster selection:
based on best stability in clustree visualization, I chose the best resolution for each sample.
```{r}
# Manually assigning resolution values to multiome_list for each sample
multiome_list$Sample_1$cluster_QC <- clustering_list_QC$Sample_1$res.0.3
multiome_list$Sample_3$cluster_QC <- clustering_list_QC$Sample_3$res.0.3
multiome_list$Sample_5$cluster_QC <- clustering_list_QC$Sample_5$res.0.3
multiome_list$Sample_7$cluster_QC <- clustering_list_QC$Sample_7$res.0.9
multiome_list$Sample_10$cluster_QC <- clustering_list_QC$Sample_10$res.0.3
multiome_list$Sample_13$cluster_QC <- clustering_list_QC$Sample_13$res.0.5
multiome_list$Sample_16$cluster_QC <- clustering_list_QC$Sample_16$res.0.9
multiome_list$Sample_18$cluster_QC <- clustering_list_QC$Sample_18$res.1.1
multiome_list$Sample_20$cluster_QC <- clustering_list_QC$Sample_20$res.0.7
multiome_list$Sample_22$cluster_QC <- clustering_list_QC$Sample_22$res.2
multiome_list$Sample_24$cluster_QC <- clustering_list_QC$Sample_24$res.1.1
multiome_list$BMMC_S1_D1$cluster_QC <- clustering_list_QC$BMMC_S1_D1$res.0.3
multiome_list$BMMC_S1_D2$cluster_QC <- clustering_list_QC$BMMC_S1_D2$res.0.3
multiome_list$BMMC_S1_D3$cluster_QC <- clustering_list_QC$BMMC_S1_D3$res.0.3
multiome_list$BMMC_S2_D1$cluster_QC <- clustering_list_QC$BMMC_S2_D1$res.0.3
multiome_list$BMMC_S3_D6$cluster_QC <- clustering_list_QC$BMMC_S3_D6$res.0.7
multiome_list$BMMC_S3_D7$cluster_QC <- clustering_list_QC$BMMC_S3_D7$res.0.9
multiome_list$BMMC_S4_D1$cluster_QC <- clustering_list_QC$BMMC_S4_D1$res.0.3
multiome_list$BMMC_S4_D8$cluster_QC <- clustering_list_QC$BMMC_S4_D8$res.0.085
multiome_list$BMMC_S4_D9$cluster_QC <- clustering_list_QC$BMMC_S4_D9$res.0.5
multiome_list$`6279DN`$cluster_QC <- clustering_list_QC$`6279DN`$res.0.5
multiome_list$`6905DN`$cluster_QC <- clustering_list_QC$`6905DN`$res.0.5


meta_data_list_qc <- list()

for (donor in names(multiome_list)) {
  meta_data <- multiome_list[[donor]]@meta.data
  meta_data_list_qc[[donor]] <- meta_data}
saveRDS(meta_data_list_qc, file = file.path(output_dir, "03_meta_data_list_11AML_9healthy_2AMS_doublet_OUTLIERS_Removed_Clusrer_QC_before_qc.rds"))


# meta_data_list_qc <- readRDS(file.path(output_dir, "03_meta_data_list_11AML_9healthy_2AMS_doublet_OUTLIERS_Removed_Clusrer_QC_before_qc.rds"))
for (donor in names(multiome_list)) {
  meta_data <- meta_data_list_qc[[donor]]
  multiome_list[[donor]]@meta.data <- meta_data}
```


# Violin plot log MAD Threshold MAD for each cluster
```{r}
nmads_up <- 3
nmads_low <- 3
log_plus_one_trans <- trans_new(
  name = "log_plus_one",
  transform = function(x) log(x + 1),
  inverse = function(x) exp(x) - 1
)

all_features <- c("nCount_RNA", "nFeature_RNA", "nCount_ATAC", "nFeature_ATAC")
group_by <- c("cluster_QC")
get_bounds <- function(data, nmads_up, nmads_low) {
    # Apply log1p transformation
    log_data <- log1p(data)
    median_log_data <- median(log_data, na.rm = TRUE)
    mad_log_data <- mad(log_data, na.rm = TRUE)

    # Calculating bounds in log scale
    lower_bound_log <- median_log_data - nmads_low * mad_log_data
    upper_bound_log <- median_log_data + nmads_up * mad_log_data

    # Convert bounds back to original scale
    lower_bound <- expm1(lower_bound_log)
    upper_bound <- expm1(upper_bound_log)

    return(c(upper_bound, lower_bound))
}


pdf(file = file.path(figures_dir, "03_violins_outliers_ddQC_4Features_log_3_3_MAD_22_samples.pdf"), width = 25, height = 22, pointsize = 5)

# A function to plot and arrange a subset of plots on a page
plot_page <- function(plots) {
  num_plots <- length(plots)
  num_cols <- 2  
  num_rows <- ceiling(num_plots / num_cols)  # Calculating the number of rows

  grid.arrange(grobs = plots, ncol = num_cols, nrow = num_rows)
}

# Iterating over all the plots
page_plots <- list()

for (sample in names(multiome_list)
     ) {
  for (group in group_by) {
    for (feature in all_features) {
     
      data <- Seurat::FetchData(multiome_list[[sample]], vars = c(feature, group))
      total_num_cells <- nrow(data)
      data[[group]] <- factor(data[[group]], levels = unique(data[[group]]))  # To ensure consistent ordering of factor levels
      outliers <- data.frame()
      clusters <- levels(data[[group]])

      outlier_counts <- vector("list", length = length(clusters))
      names(outlier_counts) <- clusters

      # Iterate over each group to calculate outliers
      for (cluster_number in clusters) {
      bounds <- get_bounds(data[[feature]][data[[group]] == cluster_number], nmads_up, nmads_low)
      feature_data_max <- bounds[1]
      feature_data_min <-  bounds[2]

      clusters_outliers <- data[(data[[feature]] < feature_data_min | data[[feature]] > feature_data_max) & data[[group]] == cluster_number, ]
      cluster_outlier_count <- nrow(clusters_outliers)
      outlier_counts[[cluster_number]] <- cluster_outlier_count
      outliers <- rbind(outliers, clusters_outliers)
        }
      max_y <- max(data[[feature]], na.rm = TRUE)
      min_y <- min(data[[feature]], na.rm = TRUE)


      data$is_outlier <- rownames(data) %in% rownames(outliers)
      # Defining colors for outliers
      outlier_colors <- c("TRUE" = "red" , "FALSE" = "#a6cee3") 
      data$outlier_color <- outlier_colors[as.character(data$is_outlier)]

       # To calculate total number of outliers for the caption
      outliers_sum <- sum(unlist(outlier_counts))
      legend_labels <- c("TRUE" = "Outlier", "FALSE" = "Passed")
      print(outliers_sum)
      try({
        plot <- ggplot(data, aes_string(x = group, y = feature)) +
          geom_jitter(aes(color = is_outlier), width = 0.3, size = 0.05, alpha = 0.8, show.legend= TRUE) +
          scale_color_manual(
          values = outlier_colors, 
          labels = legend_labels,  
          name = "QC"         
          ) +
        

         geom_violin(aes_string(color = group), trim = FALSE, alpha = 0.9, fill = NA,  show.legend= TRUE) +
         
       labs(
       title = paste0(sample, "(log+1)", ": -", nmads_low , "MAD > outliers > +", nmads_up,"MAD"),
     # subtitle = paste0("total outlier number: ",outliers_sum),
       caption = paste0("Total number of outlier detected:", outliers_sum, " out of ", total_num_cells, " cells" )
     )+

         scale_y_continuous(trans = log_plus_one_trans, breaks = round(exp(seq(log(min_y), log(max_y), by = (log(max_y) - log(min_y)) / 8))))+
         theme(panel.background = element_rect(fill = "white"), 
          plot.title = element_text(size = 10, face = "bold", hjust = 0.5,  vjust = -2),
          plot.subtitle = element_text(size = 6, color = "black",  vjust = -3, hjust = 0.5),
          plot.caption = element_text(size = 10, hjust = 0.5, color = "black"),
          axis.title = element_text(size = 6, face = "bold", color = "black"),
          axis.text = element_text(size = 6, face = "bold"))
      # Adding the plot to the list of plots
        for (cluster_level in levels(data[[group]])) {
          cluster_number <- which(levels(data[[group]]) == cluster_level)
          outlier_count <- outlier_counts[[cluster_level]]

          x_offset <- 0  
          y_multiplier <- 2.5

        # New position for text
          adjusted_x_position <- as.numeric(cluster_number) + x_offset
          adjusted_y_position <- max_y * y_multiplier
          plot <- plot + geom_text(
            aes_string(x = adjusted_x_position, y = adjusted_y_position, label = outlier_count), 
            color = "#9E0142" , 
            fill = "gray",   # Background color of the label box:  geom_label was too slow; using geom_text instead
            label.padding = unit(0.5, "lines")
            )
            }
          }, silent = TRUE)


      page_plots[[length(page_plots) + 1]] <- plot

      if (length(page_plots) == 4) {
        plot_page(page_plots)
        page_plots <- list()
      }
    }
  }
}

dev.off()

```

##Annotating outliers based on MAD 
```{r}

all_features <- c("nCount_RNA", "nFeature_RNA", "nCount_ATAC", "nFeature_ATAC")
get_bounds <- function(data, nmads_up, nmads_low) {
    # Apply log1p transformation
    log_data <- log1p(data)
    median_log_data <- median(log_data, na.rm = TRUE)
    mad_log_data <- mad(log_data, na.rm = TRUE)

    # Calculate bounds in log scale
    lower_bound_log <- median_log_data - nmads_low * mad_log_data
    upper_bound_log <- median_log_data + nmads_up * mad_log_data

    # Convert bounds back to original scale
    lower_bound <- expm1(lower_bound_log)
    upper_bound <- expm1(upper_bound_log)

    return(c(upper_bound, lower_bound))
}


OUTLIERS_FUN <- function(seurat_object, feature, clustering, nmads_up, nmads_low) {
  seurat_object@meta.data[[clustering]] <- factor(seurat_object@meta.data[[clustering]], levels = unique(seurat_object@meta.data[[clustering]]))

  all_outliers <- data.frame()  
  
  clusters <- unique(seurat_object@meta.data[[clustering]])
  
  for (cluster in clusters) {
    # Subset data based on the current cluster
    cluster_data <- seurat_object@meta.data[[feature]][seurat_object@meta.data[[clustering]] == cluster]
    
    # Calculate bounds for the current cluster
    bounds <- get_bounds(cluster_data, nmads_up, nmads_low)
    feature_data_max <- bounds[1]
    feature_data_min <- bounds[2]
    
    # Detect outliers
    clusters_outliers <- seurat_object@meta.data[(seurat_object@meta.data[[feature]] < feature_data_min | seurat_object@meta.data[[feature]] > feature_data_max) & seurat_object@meta.data[[clustering]] == cluster, ]
    all_outliers <- rbind(all_outliers, clusters_outliers)
  }
  
  # Marking outliers in seurat_object@meta.data
  outlier_rows <- rownames(seurat_object@meta.data) %in% rownames(all_outliers)
  seurat_object@meta.data$QC_MAD <- "Passed"
  seurat_object@meta.data[outlier_rows, "QC_MAD"] <- "OUTLIER"
  
  return(seurat_object)
}

  

multiome_list_outliers <- list()
for (sobj in names(multiome_list)){
  sobject <- multiome_list[[sobj]]
  for (feature_out in all_features){
  sobject <-  OUTLIERS_FUN(sobject, feature = feature_out, clustering= "cluster_QC", nmads_up= 3, nmads_low = 3)}
  multiome_list_outliers[[sobj]] <- sobject
}

for (sobj in names(multiome_list_outliers)){
  sobject <- multiome_list_outliers[[sobj]]
  print(paste0(sobj,"= ", table(sobject@meta.data$QC_MAD)))}


for (sobj in names(multiome_list_outliers)) {
  sobject <- multiome_list_outliers[[sobj]]
  qc_mad_table <- table(sobject@meta.data$QC_MAD)
  
  # To calculate the proportion of outliers
  outlier_count <- qc_mad_table["OUTLIER"]
  total_count <- sum(qc_mad_table)
  proportion_outliers <- round(outlier_count / total_count*100)
  print(paste0(sobj, " - Proportion of Outliers: ", proportion_outliers))
}

# Creating a vector to store the proportions
proportion_outliers_vector <- numeric(length(names(multiome_list_outliers)))

for (sobj in names(multiome_list_outliers)) {
  sobject <- multiome_list_outliers[[sobj]]
  qc_mad_table <- table(sobject@meta.data$QC_MAD)
  
  # Calculating the proportion of outliers
  outlier_count <- qc_mad_table["OUTLIER"]
  total_count <- sum(qc_mad_table)
  proportion_outliers <- round(outlier_count / total_count * 100)
  proportion_outliers_vector[sobj] <- proportion_outliers
  print(paste0(sobj, " - Proportion of Outliers: ", proportion_outliers))
}




# Creating a data frame with the counts of outliers and total cells for each sample
plot_data <- data.frame(
  Object = names(multiome_list_outliers),
  Total_Cells = sapply(multiome_list_outliers, function(sobj) ncol(sobj)),
  Outliers = sapply(multiome_list_outliers, function(sobj) sum(sobj@meta.data$QC_MAD == "OUTLIER"))
)

# visualisation
long_data <- gather(plot_data, key = "Category", value = "Count", -Object)

cell_colors <- c("Total_Cells" = "#386cb0", "Outliers" = "#9E0142")


# Create a stacked bar plot
p <- ggplot(long_data, aes(x = Object, y = Count, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
   scale_fill_manual(values = cell_colors) +
  labs(title = "Number of Outliers in Each Sample (Detected by MAD=+-3 on log of each cluster)",
       x = "Donors",
       y = "Cell_Count") +
  theme_minimal()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 10),  
    panel.background = element_rect(fill = "white", color = NA),  
    plot.background = element_rect(fill = "white", color = NA),   
    panel.grid = element_blank()  
  )




ggsave(file.path(figures_dir, "03_outliers_propotion_based_log_data_3_3_mad_clusters_archive.png"), width = 10, height = 6, dpi = 300)

meta_data_list_ddqc <- list()

for (donor in names(multiome_list_outliers)) {
  meta_data <- multiome_list_outliers[[donor]]@meta.data
  meta_data_list_ddqc[[donor]] <- meta_data}

saveRDS(meta_data_list_ddqc, file = file.path(output_dir, "03_meta_data_list_11AML_9healthy_2AMS_doublet_OUTLIERS_MAD3_QC_before_subseting.rds"))
multiome_list <- multiome_list_outliers
multiome_list_outliers <- NULL
# meta_data_list_ddqc <- readRDS(file.path(output_dir, "03_meta_data_list_11AML_9healthy_2AMS_doublet_OUTLIERS_MAD3_QC_before_subseting.rds"))
# for (donor in names(multiome_list)) {
#   meta_data <- meta_data_list_ddqc[[donor]]
#   multiome_list[[donor]]@meta.data <- meta_data}
```


# subsetting outliers
```{r subsetting}
table(multiome_list$BMMC_S4_D9$QC_MAD)
# OUTLIER  Passed 
#     532    5425 
final_QC_metadata_list <- list()
for (sobj_name in names(multiome_list)) {
  sobj <- multiome_list[[sobj_name]]
  print(sobj_name)
  print(length(colnames(sobj)))
  Seurat::DefaultAssay(sobj) <- "RNA"
    sobj <- subset(
      x = sobj,
      subset = QC_MAD == "Passed"
    ) 
    print(paste0("After filtering: ", length(colnames(sobj))))
    multiome_list[[sobj_name]] <- sobj
   final_QC_metadata_list[[sobj_name]] <- sobj@meta.data }
saveRDS(final_QC_metadata_list, file = file.path(output_dir, "03_meta_data_final_list_11AML_9healthy_2AMS_doublet_ddqc_all_OUTLIERS_Removed_Arcx.rds"))
```

# Cell Cycle Scoring

```{r CellCycle}
s.genes <- Seurat::cc.genes$s.genes
g2m.genes <- Seurat::cc.genes$g2m.genes
g2m.genes <- setdiff(g2m.genes, setdiff(g2m.genes, rownames(multiome_list$Sample_10)))
s.genes <- setdiff(s.genes, setdiff(s.genes, rownames(multiome_list$Sample_10)))
CELL_CYCLE <- function(x) {
  Seurat::DefaultAssay(x) <- "RNA"
  x <- NormalizeData(x)
  x <- Seurat::CellCycleScoring(x, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
    return(x)
}

multiome_list <- lapply(multiome_list, CELL_CYCLE)

table(multiome_list$Sample_10$Phase)
#  G1  G2M    S 
# 4241 1751 2224 

```

# barcode suffix correction

```{r merge}
for (donor in names(multiome_list)) {
  obj <- multiome_list[[donor]]
  obj@graphs$wknn_final <- NULL
  obj@graphs$wsnn_final <- NULL
  obj@reductions$pca_sct_final <- NULL
  obj@reductions$lsi_final <- NULL
  multiome_list[[donor]] <- obj}

colnames(multiome_list$'6279DN')<- gsub("-1", "-39", colnames(multiome_list$'6279DN'))
colnames(multiome_list$'6905DN')<- gsub("-2", "-40", colnames(multiome_list$'6905DN'))
rownames(multiome_list$'6279DN'@meta.data)
colnames(multiome_list$'6279DN'@assays$ATAC$counts)

names(multiome_list)[names(multiome_list) == "6905DN"] <- "AMS_40_6905DN"
names(multiome_list)[names(multiome_list) == "6279DN"] <- "AMS_39_6279DN"
```

#subsetting filtered cells

Subsetting the original merged object to retain only QC-passed cells identified above

```{r subsetting}
meta_data_list_final <- list()

for (donor in names(multiome_list)) {
  meta_data <- multiome_list[[donor]]@meta.data
  meta_data$CellCycle.Difference <- meta_data$S.Score - meta_data$G2M.Score
  meta_data$nCount_RNA_raw <-  meta_data$nCount_RNA
  meta_data$nFeature_RNA_raw <-  meta_data$nFeature_RNA
  meta_data$nCount_ATAC_raw <-  meta_data$nCount_ATAC
  meta_data$nFeature_ATAC_raw <-  meta_data$nFeature_ATAC 
  
  ##removing some extera columns
  meta_data <- meta_data[, !colnames(meta_data) %in% c(
  "QC_MAD", "nCount_SCT_final", "nFeature_SCT_final", "SCT_final.weight", 
  "quantile_004_percent", "nCount_SCT_QC", "nFeature_SCT_QC", "SCT_QC.weight", 
  "cluster_QC", "nCount_SCT_doub", "nFeature_SCT_doub", "SCT_doub.weight", 
  "scDblFinder_RNA_counts", "scDblFinder_RNA_counts_score", 
  "scDblFinder_ATAC_counts", "scDblFinder_ATAC_counts_score", "old.ident", "cluster_doublet", "ATAC.weight",
  "nCount_RNA","nFeature_RNA", "nCount_ATAC", "nFeature_ATAC")]


   multiome_list[[donor]]@meta.data <- meta_data
  meta_data_list_final[[donor]] <- meta_data}

#merging all the barcodes from different objects
  barcode_list <- c()
  for (obj in names(meta_data_list_final)){
    barcode_list <- c(barcode_list , rownames(meta_data_list_final[[obj]]))
    }
 length(barcode_list) #138877
 #I removed 27446 doublet and the rest during removing outliers

saveRDS(meta_data_list_final, file = file.path(output_dir, "03_final_meta_data_list_11AML_9healthy_2AMS_after_qc_main_clusters_annotated.rds"))
```

# merging metadatas
I only merged meta.data dfs from different filtered objs.

```{r merging}
# Combine the list of dataframes into one large dataframe while retaining row names
merged_meta_data_22 <- bind_rows(meta_data_list_final) 

dim(merged_meta_data_22) #138877     99
saveRDS(merged_meta_data_22, file = file.path(output_dir, "03_final_merged_meta_data_11AML_9healthy_2AMS_after_qc_main_clusters_annotated_barcode39_40_v2.rds"))

```

# Subsetting merged data
```{r subsetting data}

Merge_obj  <- readRDS(file.path(output_dir, "01_merged_multiome_11AML_9_healthy_2AMS_before_doublet_and_qc.rds"))
dim(Merge_obj) #184466 185243
merged_meta_data <- readRDS(file.path(output_dir, "03_final_merged_meta_data_11AML_9healthy_2AMS_after_qc_main_clusters_annotated_barcode39_40_v2.rds"))
dim(merged_meta_data) #138877     99


filtered_barcodes <- rownames(merged_meta_data)
length(filtered_barcodes) #138877

Idents(Merge_obj) <- colnames(Merge_obj)


# subsetting
Merge_obj <- subset(
  x = Merge_obj,
  idents = filtered_barcodes
)
dim(Merge_obj) #184466 138877

identical(
rownames(Merge_obj@meta.data),
rownames(merged_meta_data)) #FALSE

setdiff(
  rownames(Merge_obj@meta.data),
rownames(merged_meta_data))

 
# Reorder merged_meta_data to match the rownames of Merge_obj@meta.data
merged_meta_data <- merged_meta_data[rownames(Merge_obj@meta.data), ]


identical(
rownames(Merge_obj@meta.data),
rownames(merged_meta_data)) #TRUE


identical(
Merge_obj@meta.data$sample_id,
merged_meta_data$sample_id) #TRUE

identical(
Merge_obj@meta.data$barcode,
merged_meta_data$barcode) #FALSE

Merge_obj@meta.data <- merged_meta_data
dim(Merge_obj) #184466 138877
dim(Merge_obj@meta.data) #138877     99

Merge_obj@meta.data$barcode_10x <- Merge_obj@meta.data$barcode
Merge_obj@meta.data$barcode <- colnames(Merge_obj)

identical(merged_meta_data$barcode, Merge_obj@meta.data$barcode_10x) #TRUE
identical(colnames(Merge_obj), Merge_obj@meta.data$barcode) #TRUE
identical(colnames(Merge_obj), as.character(Idents(Merge_obj))) #TRUE


```

# Converting discrete data to Categorical for later visualisation 
```{r Categorical, fig.width=30, fig.height=10}
columns_to_process <- c("nCount_RNA_raw", "nFeature_RNA_raw", "nFeature_ATAC_raw", "nCount_ATAC_raw", 
                        "percentage_chr7_peaks", "percentage_chr7_genes", 
                        "percentage_RB", "percentage_Malat1", "percentage_HB", 
                        "percentage_MT", "percentage_FLT3", "percentage_t_cells",
                        "percentage_B_cell_markers","percentage_monocytes", 
                        "percentage_megakaryocytes", "percentage_erythroid_cells", "percentage_chrY_peaks", "percentage_chr21_peaks", "percentage_chr15_peaks")

for (column in columns_to_process) {
  new_col <- paste0(column, "_quantile_merged")

  Merge_obj@meta.data <- Merge_obj@meta.data %>%
    mutate(!!new_col := ntile(!!sym(column), 10)) %>%
    mutate(!!new_col := case_when(
      !!sym(new_col) == 1 ~ "Q1",
      !!sym(new_col) == 2 ~ "Q2",
      !!sym(new_col) == 3 ~ "Q3",
      !!sym(new_col) == 4 ~ "Q4",
      !!sym(new_col) == 5 ~ "Q5",
      !!sym(new_col) == 6 ~ "Q6",
      !!sym(new_col) == 7 ~ "Q7",
      !!sym(new_col) == 8 ~ "Q8",
      !!sym(new_col) == 9 ~ "Q9",
      !!sym(new_col) == 10 ~ "Q10"
    ))
}

Merge_obj@meta.data <- Merge_obj@meta.data %>%
  mutate(Monosomy_cutoff =  ifelse(Merge_obj@meta.data$percentage_chr7_peaks < 3 , "Under_3", ifelse(Merge_obj@meta.data$percentage_chr7_peaks > 3.7 , "above_3.7", "3-3.7")))
  
  #          AML_monosomy AML_non_monosomy healthy_donors
  # 3-3.7              871             1247            903
  # above_3.7         7558            55257          55766
  # Under_3          16951              205            119

table(Merge_obj$Monosomy_cutoff,Merge_obj$health_condition )
table(Merge_obj$health_condition)
  # AML_monosomy AML_non_monosomy   healthy_donors 
   #           25380            56709            56788 

##############################
Merge_obj$donor_main_clusters <- paste0(Merge_obj$sample_id,"_" , Merge_obj$cluster_5_main_per_sample)
length(unique(Merge_obj$donor_main_clusters)) #113
```

# Adding cell type annotations from public dataset
```{r}
meta_data_22 <- Merge_obj@meta.data
meta_data_22$barcode <- rownames(meta_data_22)
meta_data_lueken <- read.csv(file.path(data_dir, "adata_obs.csv"))
dim(meta_data_lueken) # 69249    29
dim(meta_data_22) #138877    127
colnames(meta_data_lueken)
unique(meta_data_lueken$batch)
unique(meta_data_22$donor)
table(meta_data_lueken$cell_type, useNA = "ifany")

# Create a mapping
batch_mapping <- c(
  "s1d1"  = "BMMC_S1_D1",
  "s1d2"  = "BMMC_S1_D2",
  "s1d3"  = "BMMC_S1_D3",
  "s2d1"  = "BMMC_S2_D1",
  "s2d4"  = "BMMC_S2_D4",  
  "s2d5"  = "BMMC_S2_D5", 
  "s3d3"  = "BMMC_S3_D3",  
  "s3d6"  = "BMMC_S3_D6",
  "s3d7"  = "BMMC_S3_D7",
  "s3d10" = "BMMC_S3_D10", 
  "s4d1"  = "BMMC_S4_D1",
  "s4d8"  = "BMMC_S4_D8",
  "s4d9"  = "BMMC_S4_D9"
)

meta_data_lueken <- meta_data_lueken %>%
  mutate(donor = batch_mapping[batch])

unique(meta_data_lueken$donor)

rownames(meta_data_22)
metadata <- meta_data_22

# Extracting the suffix using regular expressions
metadata <- metadata %>%
  mutate(suffix = sub(".*-(\\d+)$", "\\1", rownames(metadata)))
unique(metadata$suffix)
identical(metadata$suffix, metadata$sample_id) #TRUE

unique(meta_data_lueken$donor)
unique(metadata$donor)

meta_data_lueken_subset <- meta_data_lueken %>%
 dplyr::filter(donor %in% metadata$donor)
dim(meta_data_lueken_subset) #47137    30
table(metadata$health_condition) #  56788 healthy

unique(meta_data_lueken_subset$donor)

table(metadata$donor, metadata$sample_id)
table(meta_data_lueken_subset$donor)
# left_join using dplyr
meta_data_lueken_subset$donor <- unname(meta_data_lueken_subset$donor)

suffix_mapping <- c(
"BMMC_S1_D1"="26",
"BMMC_S1_D2"="27",
"BMMC_S1_D3"="28",
"BMMC_S2_D1"="29",
"BMMC_S3_D6"="34",
"BMMC_S3_D7"="35",
"BMMC_S4_D1"="36",
"BMMC_S4_D8"="37",
"BMMC_S4_D9"="38")

meta_data_lueken_2 <- meta_data_lueken_subset %>%
  mutate(suffix = suffix_mapping[donor])

replace_suffix <- function(x, old_suffix, new_suffix) {
  sub(paste0("-1-", old_suffix), new_suffix, x)
}

meta_data_lueken_2$X <- mapply(replace_suffix, meta_data_lueken_2$X, meta_data_lueken_2$suffix, paste0("-", meta_data_lueken_2$suffix))
head(meta_data_lueken_2$X)
unique(meta_data_lueken_2$suffix)

meta_data_lueken_2$suffix <- unname(meta_data_lueken_2$suffix)

meta_data_lueken_3 <- meta_data_lueken_2 %>%
  mutate(barcode = mapply(function(x, suffix) sub("-[0-9]+-.*", paste0("-", suffix), x), X, suffix))
head(meta_data_lueken_3$barcode)


meta_data_lueken_3$barcode <- unname(meta_data_lueken_3$barcode)
dim(meta_data_lueken_3) #47137    32


meta_data_lueken_subset <- meta_data_lueken_3 %>%
 dplyr::filter(barcode %in% metadata$barcode)
dim(meta_data_lueken_subset) #23964    128
dim(metadata) #115991    128

merged_data <- metadata %>%
  left_join(meta_data_lueken_subset, by = "barcode")

dim(merged_data) #115991    159
table(merged_data$cell_type)

sum(is.na(merged_data$cell_type))#96982
table(merged_data$cell_type, useNA = "ifany")

 # B1 B          CD14+ Mono          CD16+ Mono    CD4+ T activated        CD4+ T naive              CD8+ T        CD8+ T naive 
 #               1310                4915                 896                3836                3282                8758                 382 
 #               cDC2        Erythroblast            G/M prog                 HSC ID2-hi myeloid prog                 ILC          Lymph prog 
 #                409                2551                 519                 442                 100                 454                1016 
 #          MK/E prog       Naive CD20+ B                  NK          Normoblast                 pDC         Plasma cell     Proerythroblast 
 #                536                3316                4444                 878                 466                 276                1167 
 #     Transitional B                <NA> 
 #               1942               96982 
 #     
     
rownames(merged_data) <- merged_data$barcode

identical(rownames(Merge_obj@meta.data), rownames(merged_data)) #TRUE

Merge_obj@meta.data$cell_type_Lucken <- merged_data$cell_type
Merge_obj@meta.data$QCMeds_Lucken <- merged_data$QCMeds
Merge_obj@meta.data$ATAC_pseudotime_order_Lucken <- merged_data$ATAC_pseudotime_order
Merge_obj@meta.data$GEX_pseudotime_order_Lucken <- merged_data$GEX_pseudotime_order
Merge_obj@meta.data$GEX_phase_Lucken <- merged_data$GEX_phase


saveRDS(Merge_obj, file = file.path(output_dir, "03_merged_object_11AML_9healthy_2AMS_with_merged_meta_data_after_qc_all_features_included_v2.rds")) ##final
```

# Removing sample-specific features

Genes and ATAC peaks from chromosomes X, Y, and 7 are removed, along with immunoglobulin/TCR 
variable genes, haemoglobin (HB), ribosomal (RB), and mitochondrial (MT) genes, 
as these are confounded by sample-specific biology.
```{r features, fig.width=30, fig.height=20}
DefaultAssay(Merge_obj) <- "RNA"
ensembl <- useMart("ensembl", dataset = "hsapiens_gene_ensembl" )
pages = attributePages(ensembl)
filters = listAttributes(ensembl)

#########  chrX
chrX_genes <- getBM(
  attributes = c('hgnc_symbol', 'chromosome_name', 'start_position','end_position'),
  filters = 'chromosome_name',
  values = 'X',
  mart = ensembl
)

chrX_genes <- chrX_genes$hgnc_symbol
length(chrX_genes) #2496

########## chr Y
chrY_genes <- getBM(
  attributes = c('hgnc_symbol', 'chromosome_name', 'start_position','end_position'),
  filters = 'chromosome_name',
  values = 'Y',
  mart = ensembl
)

chrY_genes <- chrY_genes$hgnc_symbol
length(chrY_genes) #603

######## chr7
chr7_genes <- getBM(
  attributes = c('hgnc_symbol', 'chromosome_name'),
  filters = 'chromosome_name',
  values = '7',
  mart = ensembl
)
chr7_genes <- chr7_genes$hgnc_symbol
length(chr7_genes) # 1984
chr7_genes <- chr7_genes[chr7_genes%in%rownames(Merge_obj)]
length(chr7_genes) # 1062


################## 
all_genes_name <- rownames(Merge_obj@assays$RNA)
length(all_genes_name) #36601
removing_genes_patterns <- "^IGHV|^IGKV|^IGLV|^TRAV|^TRBV|^TRDV|^TRGV|^HB[^(P)]|MALAT1|^RP[SL]|^IG[A-Z]*|^MT-"
removing_genes <- grep(removing_genes_patterns, all_genes_name, value = TRUE)

length(removing_genes) #726
genes_to_remove <- unique(c(chrX_genes, chrY_genes, chr7_genes, removing_genes))
length(genes_to_remove) #4018

###############################

gene_coords <- getBM(attributes = c('hgnc_symbol', 'chromosome_name', 'start_position', 'end_position'), 
                     filters = 'hgnc_symbol', values = genes_to_remove, mart = ensembl)
dim(gene_coords) #4539    4
gr_genes_to_remove <- GRanges(gene_coords$chromosome_name, IRanges(gene_coords$start_position, gene_coords$end_position))

seqlevelsStyle(gr_genes_to_remove) <- "UCSC"
all_peaks <- rownames(Merge_obj@assays$ATAC)
length(all_peaks) #184466
atac_peaks <- StringToGRanges(all_peaks, sep = c("-", "-"))
length(atac_peaks) #184466

promoter_extension <- 100000

extend_ranges <- function(ranges, strand) {
  new_start <- ifelse(strand == "+", 
                      start(ranges) - promoter_extension, 
                      start(ranges))
  new_end <- ifelse(strand == "-", 
                    end(ranges) + promoter_extension, 
                    end(ranges))

  new_start <- pmax(new_start, 1)  # Ensure start is not less than 1
  IRanges(new_start, new_end)
}

# Initialize a list to store extended ranges
extended_ranges_list <- vector("list", length(gr_genes_to_remove))

# Loop over the indices of gr_genes_to_remove and extend ranges
for (i in seq_along(gr_genes_to_remove)) {
  extended_ranges_list[[i]] <- extend_ranges(ranges(gr_genes_to_remove)[i],
                                             strand(gr_genes_to_remove)[i])
}

# Combine the extended ranges into a single IRanges object
combined_extended_ranges <- do.call("c", extended_ranges_list)
length(combined_extended_ranges) #4539

# Update the GRanges object with extended ranges
gr_genes_to_remove <- GRanges(seqnames = seqnames(gr_genes_to_remove),
                              ranges = combined_extended_ranges,
                              strand = strand(gr_genes_to_remove))

# Find overlapping peaks
overlaps <- findOverlaps(atac_peaks, gr_genes_to_remove) 
length(overlaps) #10161
# Remove overlapping peaks
filtered_atac_peaks <- atac_peaks[-queryHits(overlaps),]

length(filtered_atac_peaks) #175006

# Proceed the analysis using filtered_atac_peaks


chr <- as.character(seqnames(filtered_atac_peaks))
unique(chr)
start <- start(filtered_atac_peaks)
end <- end(filtered_atac_peaks)

filtered_atac_peaks_strings <- paste(chr, start, end, sep = "-")
length(intersect(filtered_atac_peaks_strings,all_peaks)) #175006
# To filter strings that start with chrX or chrY
residual_peaks <- filtered_atac_peaks_strings[-grep("^chrX|^chrY|^chr7", filtered_atac_peaks_strings)]
length(residual_peaks) #170327
# Check specific patterns individually
matches_chrX <- grep("^chrX", filtered_atac_peaks_strings, value = TRUE)
matches_chrY <- grep("^chrY", filtered_atac_peaks_strings, value = TRUE)
matches_chr7 <- grep("^chr7", filtered_atac_peaks_strings, value = TRUE)

print(length(matches_chrX))  # Expecting 0 based on the previous results 1759
print(length(matches_chrY))  # Expecting 0 based on the previous results 72
print(length(matches_chr7))  # Expecting non-zero, as it matched before 2848


matches <- grep("^chrX|^chrY|^chr7", filtered_atac_peaks_strings)
length(matches) #4679
# Exclude the matches properly
if (length(matches) > 0) {
    residual_peaks <- filtered_atac_peaks_strings[-matches]
} else {
    residual_peaks <- filtered_atac_peaks_strings
}

length(residual_peaks) #170327
length(intersect(all_genes_name,genes_to_remove)) #2708

residual_genes <- all_genes_name[!all_genes_name %in% genes_to_remove]
length(residual_genes) #33893

residual_features <- c(residual_peaks, residual_genes)
length(residual_features) #204220
Merge_obj #221067 features across 138877 samples within 2 assays 
Merge_obj <- subset(Merge_obj, features = residual_features) #204220 features across 138877 samples within 2 assays 
saveRDS(Merge_obj, file = file.path(output_dir, "03_merged_object_11AML_9healthy_2AMS_with_merged_meta_data_after_qc_filtered_features_ch7_removed.rds")) #final
```



