---
title: "01_02_PreProcessing_SoupX_Creating_Multiome_Sobj"
author: "Reza Ghamsari"
date: "2023-11-10"
output: html_document
---

```{r setup, include=FALSE}
# Store execution times for each chunk
all_times <- list()

# Set custom knitr hook to track time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()  # Start timer
    } else {
      res <- difftime(Sys.time(), now)  # Calculate elapsed time
      all_times[[options$label]] <<- res  # Store time for the chunk
    }
  }
}))

# Set global chunk options
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  time_it = TRUE  # Enable time tracking
)


```

#Directory_paths
```{r}
my_colors <-c("#9d183299","#00A087FF", "#525ecc99","#E9967A", "#4DBBD5FF",  "#003C67FF", "#386cb0", "#7E6148FF",  "#ADB6B6FF", "#52a7cc99", "#F0E685FF", "#cc52c099",  "#6acc5299", "#802268FF" ,"#ffbb78", "#3C548899", "#189b3699", "#cc9b5299", "#1f0fdf99", "#ccebc5", "#cc527699", "#42857599", "#868686FF", "#006400", "#0A47FFFF" , "#3B1B53FF",  "#749B58FF" ,"#ccc05299", "#cc765299" , "#1B1919FF", "#00D68FFF" ,"#14FFB1FF","#3C5488FF" , "#8F7700FF" , "#164194FF" , "#0094CDFF" , "#FFDAB9", "#FF00FF","#00FFFF","green" , "blue",  "#DC143C","yellow")

# Define the directory to save the PDF files
fig_path_stnx <- "/home/users/allstaff/ghamsari.r/seurat_figures/MiniAtlas/"
# Define the directory to save and load the RDS files
rds_path_vst <- "/vast/projects/aml_multiome/RaCHseq_AML/seurat_outputs/MiniAtlas/rds_files/"
# Path to the directory of raw data
path_to_AML <- "/stornext/General/data/user_managed/grpu_mritchie_1/reza/projects/mat_seq/RaCHseq_AML/cellranger-arc_outputs/count_outputs/"
path_to_BMMC <- "/stornext/General/data/user_managed/grpu_mritchie_1/reza/projects/mat_seq/BMMC_luecken/cellranger-arc_outputs/count_outputs/"
```
**Script Goal:**
The primary goal of this project is to curate a concise atlas comprising 25 AML (Acute Myeloid Leukemia) samples, while integrating data from four healthy Bone Marrow Mononuclear Cells (BMMCs) obtained from publicly available sources.
I merged all the 25AML and 13 healthy samples (**script: 00_02_03_cellranger_arc_aggr_38_AML_Public_ATAC_GEX.sh**) and then subsampled these 29 samples from in python (**script: 01_01_Sub_sampeling_29_donors.ipynb**) to analyse them in R.

**Script Summary:**

Soupx Integration:

The script estimates ambient RNA contamination using SoupX . For SoupX, raw data is required, so the Cell Ranger ARC count output is used directly rather than the aggregated data. The raw RNA counts are then processed to estimate and subtract the ambient RNA contributions for each sample.


Clustering and Resolution Selection:

The optimal clustering resolution is chosen for each sample based on the stability observed in Clustree visualizations. This clustering process uses default highly variable genes (HVGs) for accurate cluster identification.

Ambient RNA Visualisation:

The script calculates the percentage of ambient RNA in each sample by comparing the total RNA counts with the ambient RNA counts (using SoupX). The results are stored in a data frame and saved as a CSV file. Additionally, a stacked bar plot is generated to visualize the ambient RNA vs. corrected RNA counts for each sample.

A list of Seurat objects is created by converting each SoupX-adjusted matrix into a Seurat object. Then this list merged to create one seurat object.

Reading and Processing ATAC Data:

The ATAC data, originally generated by Cell Ranger ARC -aggr, is processed and subsampled into a .mtx format for 25 patient samples and 4 healthy donor samples.

The ATAC data is then converted into a ChromatinAssay object, and is added to assays slot in seurat object. At the end of this script we will have one Multiome_object containing both ATAC and Corrected RNA assays.



#Loading Packages
```{r, pkgs}
script_title <- "01_02_PreProcessing_SoupX_Creating_Multiome_Sobj"
library(Seurat, lib.loc = "/home/users/allstaff/ghamsari.r/R/x86_64-pc-linux-gnu-library/old_versions")
packageVersion("Seurat") #used  the ‘4.4.0’ version for ATAC data integration
library(Signac)
library(ggplot2)
library(SoupX)
library(scCustomize)
library(gridExtra)
library(parallel)
library(clustree)
library(RColorBrewer)
library(patchwork)
library(igraph)
library(tibble)
library(ddqcR)
library(unix)
library(gtools)
library(tidyr)
library(dplyr)
library(tidyverse)
sink(paste0(fig_path_stnx,script_title ,"_sessionInfo.txt"))
sessionInfo()
sink()

script_title
```

#SLURM Memory Management
This section ensures smooth execution in a shared computing environment by managing SLURM memory limits, particularly addressing issues with forking and parallelism in SLURM when using R. It checks the available memory and adjusts settings accordingly.
```{r,}
# Get available memory in bytes (Linux command)
system_info <- system("free -b | grep Mem", intern = TRUE)
mem_info <- strsplit(system_info, " +")[[1]]

# Extract available and total memory
available_ram <- as.numeric(mem_info[4])  # Available RAM in bytes
total_memory_bytes <- as.numeric(mem_info[2])  # Total memory

# Available memory (free + buffers/cache)
available_memory_bytes <- as.numeric(mem_info[6])

# Print memory information
print(paste("Available RAM (bytes):", available_ram * 2^20))  # Convert to MB
print(paste("Total memory (bytes):", total_memory_bytes))
print(paste("Available memory (bytes):", available_memory_bytes))

# Retrieve SLURM memory limit for the node
slurm_mem_available_B <- as.numeric(Sys.getenv("SLURM_MEM_PER_NODE", unset = NA)) * 10^6

# Print SLURM memory limit
print(paste("SLURM memory limit:", slurm_mem_available_B))
# Set memory usage limit, keeping 2GB safe margin
Memory_to_use <- slurm_mem_available_B - (2000 * 2^20)  # 2000 MB = 2 GB

# Set both hard and soft memory limits
unix::rlimit_as(Memory_to_use, Memory_to_use)

# Limit the core dump size to 55MB to prevent large core dumps on crash
rlimit_core(cur = 55, max = 55)

```
#Loading GEX Data
To perform ambient RNA detection, I extracted the raw RNA data from each individual sample's Cell Ranger Arc count output(**script: 00_01_03_cellranger_arc_count_13_public_data_ATAC_GEX.sh & 00_02_02_cellranger_arc_count_25AML_ATAC_GEX.sh**), rather than from the merged dataset generated with Cell Ranger Arc aggr (**script: 00_02_03_cellranger_arc_aggr_38_AML_Public_ATAC_GEX.sh**). Later, I will integrate this with the merged ATAC-seq dataset.
```{r,}
# Define the base directories (25 AML and 4 healthy)
base_dirs <- c(rep(path_to_AML, 25),
               rep(path_to_BMMC, 4))
# List all directories (sample names) from the AML and BMMC directories
list_of_AML_samples <- list.dirs(path_to_AML, recursive = FALSE, full.names = FALSE)
list_of_BMMC_samples <- list.dirs(path_to_BMMC, recursive = FALSE, full.names = FALSE)
# Order the AML samples alphabetically
list_of_AML_samples <- list_of_AML_samples[mixedorder(list_of_AML_samples)]

# Print the ordered list
print(list_of_BMMC_samples)
print(list_of_AML_samples)

# Concatenate the lists
sample_ids <- c(list_of_AML_samples, list_of_BMMC_samples)
print(sample_ids)
length(sample_ids) #38
# Create a vector with numbers 1 through 25
AML_samples_numbers <- seq(1, 25)
# Additional numbers to add
Healthy_sample_numbers <- c(26, 29, 36, 37)

# Combine the two vectors
sample_ids <- sample_ids[c(AML_samples_numbers, Healthy_sample_numbers)]

length(sample_ids) #29

# Initialize list to store Seurat objects
seurat_list <- list()



# Create Seurat objects for each sample
for (i in seq_along(sample_ids)) {
  sample_id <- sample_ids[i]
  base_dir <- base_dirs[i]

  # Read the data
  sample <- Read10X_h5(paste0(base_dir, sample_id, "/outs/filtered_feature_bc_matrix.h5"))

  # Extract RNA and ATAC data
  rna_counts <- sample$`Gene Expression`

  # Create Seurat object
  seurat_obj <- CreateSeuratObject(counts = rna_counts)

  
  # Add Seurat object to list
  seurat_list[[sample_id]] <- seurat_obj
}


for (id in names(seurat_list)) {
  seurat_list[[id]]$sample_id <- id
}


for (id in names(seurat_list)){
Idents(seurat_list[[id]]) <- id
}
```
#Clustering 
As expression data is highly sparse at the single cell level, it is highly recommended that clustering
information be provided for ambient RNA detection to allow the subtraction method to share information between cells. Without
grouping cells into clusters, it is difficult (and usually impossible) to tell the difference between
a count of 1 due to background contamination and a count of 1 due to endogenous expression. This
ambiguity is removed at the cluster level where counts can be aggregated across cells. This information
can then be propagated back to the individual cell level to provide a more accurate removal
of contaminating counts.

```{r, clustering}
# Check if the raw counts and normalized data are identical before normalizing
identical(seurat_list$Sample_1@assays$RNA$counts, seurat_list$Sample_1@assays$RNA$data) #TRUE
# Define a custom normalization function
NORMAL_LIST <- function(x) {
    DefaultAssay(x) <- "RNA"
    x <- NormalizeData(x,  assay = "RNA")
     return(x)}
# Use parallel processing with mclapply to normalize data for all Seurat objects
seurat_list <- mclapply(seurat_list, NORMAL_LIST, mc.cores =  detectCores()-5)
# Verify that normalization has been applied (counts and normalized data should be different now)
identical(seurat_list$Sample_1@assays$RNA$counts, seurat_list$Sample_1@assays$RNA$data) #FALSE
FIND_HVG_LIST_VST_CLUSTER <- function(x) {
  DefaultAssay(x) <- "RNA"  # Set RNA as the default assay
  
  # Find highly variable genes (HVG) using the VST method
  x <- FindVariableFeatures(x, nfeatures = 2000, selection.method = "vst")
  
  # Scale the data for downstream analysis
  x <- ScaleData(x, assay = "RNA")
  
  # Perform PCA for dimensionality reduction
  x <- RunPCA(x, reduction.name = 'pca_vst_default_raw')
  
  # Perform UMAP for visualization
  x <- RunUMAP(x, reduction = 'pca_vst_default_raw', dims = 1:50, reduction.name = 'umap.vst', reduction.key = 'vstUMAP_')
  
  # Find neighbors based on PCA for clustering
  x <- FindNeighbors(x, reduction = 'pca_vst_default_raw', dims = 1:50, graph.name = 'graph_vst_default_raw')
  
  # Perform clustering using a resolution range to capture different cluster densities
  x <- FindClusters(x, graph.name = "graph_vst_default_raw", algorithm = 3, verbose = FALSE, resolution = seq(0.2, 1.5, 0.05))
  
  # Store cluster labels in metadata for later use
  x@meta.data$clusters_vst_default_raw <- x@meta.data$seurat_clusters
  
  return(x)
}

seurat_list <- mclapply(seurat_list, FIND_HVG_LIST_VST_CLUSTER, mc.cores = detectCores() - 5, mc.preschedule = FALSE)

```

##Clustree
```{r, clustree ovelay, fig.width=20}
# Function to generate Clustree plots for a list of Seurat objects
CLUSTREE_PLOT <- function(seurat_list, fig_path, my_colors) {

  # Open a PDF device to save the plot as a multi-page document
  pdf(file = paste0(fig_path_stnx, "01_02_CLUSTREE_before_soupx.pdf"), width = 30, height = 25, pointsize = 5) # Increase pointsize for better readability

  # Function to arrange plots into a grid
  plot_page <- function(plots) {
    num_plots <- length(plots)
    num_cols <- 2  # Number of columns in the grid
    num_rows <- ceiling(num_plots / num_cols)  # Calculate the number of rows
    
    # Arrange and display the plots in a grid layout
    grid.arrange(grobs = plots, ncol = num_cols, nrow = num_rows)
  }
  
  # Initialize a list to store individual plots for each page
  page_plots <- list()

  # Iterate through all the Seurat objects in the provided list
  for (obj in names(seurat_list)) {
    sobj <- seurat_list[[obj]]
    
    # Iterate over different clustering prefixes
       prefix <- "graph_vst_default_raw_res." 
      
      # Generate the Clustree plot for each Seurat object and clustering prefix
      plot <- clustree(sobj, prefix = prefix, node_alpha = 0.5) + 
        scale_color_manual(values = my_colors) + 
        guides(color = guide_legend(ncol = 2)) + 
        ggtitle(paste("Clustree", obj, prefix))
      
      # Add the plot to the list of page plots
      page_plots[[length(page_plots) + 1]] <- plot

      # Once 2 plots are accumulated, arrange them into a page
      if (length(page_plots) == 2) {
        plot_page(page_plots)
        page_plots <- list()  # Reset the list for the next set of plots
      
    }
  }

  # Close the PDF device to save the plots to the file
  dev.off()
}

# Call the function with the Seurat list, output file path, and color palette
CLUSTREE_PLOT(seurat_list, fig_path_stnx, my_colors)
```

#Choosing the Resolution
The optimal resolution for each sample was selected based on the stability observed in the Clustree visualization. 
```{r,}
seurat_list$Sample_1$cluster_soupx <- seurat_list$Sample_1$graph_vst_default_raw_res.1
seurat_list$Sample_2$cluster_soupx <- seurat_list$Sample_2$graph_vst_default_raw_res.1
seurat_list$Sample_3$cluster_soupx <- seurat_list$Sample_3$graph_vst_default_raw_res.0.8
seurat_list$Sample_4$cluster_soupx <- seurat_list$Sample_4$graph_vst_default_raw_res.1.15 #very unstable
seurat_list$Sample_5$cluster_soupx <- seurat_list$Sample_5$graph_vst_default_raw_res.0.5
seurat_list$Sample_6$cluster_soupx <- seurat_list$Sample_6$graph_vst_default_raw_res.1
seurat_list$Sample_7$cluster_soupx <- seurat_list$Sample_7$graph_vst_default_raw_res.0.9
seurat_list$Sample_8$cluster_soupx <- seurat_list$Sample_8$graph_vst_default_raw_res.0.4
seurat_list$Sample_9$cluster_soupx <- seurat_list$Sample_9$graph_vst_default_raw_res.0.7
seurat_list$Sample_10$cluster_soupx <- seurat_list$Sample_10$graph_vst_default_raw_res.1
seurat_list$Sample_11$cluster_soupx <- seurat_list$Sample_11$graph_vst_default_raw_res.0.5
seurat_list$Sample_12$cluster_soupx <- seurat_list$Sample_12$graph_vst_default_raw_res.0.85
seurat_list$Sample_13$cluster_soupx <- seurat_list$Sample_13$graph_vst_default_raw_res.0.7
seurat_list$Sample_14$cluster_soupx <- seurat_list$Sample_14$graph_vst_default_raw_res.0.5
seurat_list$Sample_15$cluster_soupx <- seurat_list$Sample_15$graph_vst_default_raw_res.1
seurat_list$Sample_16$cluster_soupx <- seurat_list$Sample_16$graph_vst_default_raw_res.0.8
seurat_list$Sample_17$cluster_soupx <- seurat_list$Sample_17$graph_vst_default_raw_res.0.9
seurat_list$Sample_18$cluster_soupx <- seurat_list$Sample_18$graph_vst_default_raw_res.0.9
seurat_list$Sample_19$cluster_soupx <- seurat_list$Sample_19$graph_vst_default_raw_res.1.15
seurat_list$Sample_20$cluster_soupx <- seurat_list$Sample_20$graph_vst_default_raw_res.0.85
seurat_list$Sample_21$cluster_soupx <- seurat_list$Sample_21$graph_vst_default_raw_res.1.15
seurat_list$Sample_22$cluster_soupx <- seurat_list$Sample_22$graph_vst_default_raw_res.0.6
seurat_list$Sample_23$cluster_soupx <- seurat_list$Sample_23$graph_vst_default_raw_res.0.75
seurat_list$Sample_24$cluster_soupx <- seurat_list$Sample_24$graph_vst_default_raw_res.0.85
seurat_list$Sample_25$cluster_soupx <- seurat_list$Sample_25$graph_vst_default_raw_res.1.2
seurat_list$BMMC_S1_D1$cluster_soupx <- seurat_list$BMMC_S1_D1$graph_vst_default_raw_res.0.9
seurat_list$BMMC_S2_D1$cluster_soupx <- seurat_list$BMMC_S2_D1$graph_vst_default_raw_res.1.25
seurat_list$BMMC_S4_D1$cluster_soupx <- seurat_list$BMMC_S4_D1$graph_vst_default_raw_res.0.6
seurat_list$BMMC_S4_D8$cluster_soupx <- seurat_list$BMMC_S4_D8$graph_vst_default_raw_res.0.5

```

#Soupx 
SoupX did not work with object Seurat v5, so I ended up loading V4 and also reading filtered Data again.
https://rawcdn.githack.com/constantAmateur/SoupX/204b602418df12e9fdb4b68775a8b486c6504fe4/inst/doc/pbmcTutorial.html
```{r, soupX , fig.width=20}
# Initialize list to store SoupChannel objects
soupx_list <- list()
sc_list <- list()
rna_count_list <- list()
# Create Seurat objects for each sample
for (i in seq_along(seurat_list)) {
  base_dir <- base_dirs[i]
  Obj_name <- names(seurat_list)[i]
  sobj <- seurat_list[[Obj_name]]
    # Read the data
  # sample <- Read10X_h5(paste0(base_dir, Obj_name, "/outs/filtered_feature_bc_matrix.h5"))
  # Extract RNA and ATAC data
  # rna_counts<- sample$`Gene Expression`
  rna_counts <- sobj@assays$RNA@counts
  # Read the data
  raw <- Read10X(paste0(base_dir, Obj_name, "/outs/raw_feature_bc_matrix"))
  raw <- raw$`Gene Expression`
  sc <- SoupChannel(raw,rna_counts)
  sc <- setClusters(sc,sobj@meta.data$cluster_soupx)
  #A vector of cluster IDs, named by cellIDs. If NULL clusters auto-loaded from sc. If FALSE, no clusters are used.
  #The idea of this method is that genes that are highly expressed in the soup and are marker genes for
# some population can be used to estimate the background contamination. Marker genes are identified
# using the tfidf method (see quickMarkers). The contamination fraction is then calculated at the
# cluster level for each of these genes and clusters are then aggressively pruned to remove those that
# give implausible estimates.
  sc <- autoEstCont(sc, doPlot=FALSE)
  out <- adjustCounts(sc, roundToInt = TRUE)
  # Add to list with sample id as name
  soupx_list[[Obj_name]] <- out
  }


```
#exporting csv results
```{r,}
#saving the percentage of each samples' ambient RNA in a data frame
results <- data.frame()
for (obj in names(seurat_list)) {
  sobj <- seurat_list[[obj]]
  DefaultAssay(sobj) <- "RNA"
  rna_counts <- sobj@assays$RNA@counts
   # Check if column and row names match between the Seurat and SoupX data
  if (identical(colnames(soupx_list[[obj]]), colnames(rna_counts)) &
      identical(rownames(soupx_list[[obj]]), rownames(rna_counts))) {
    Total_Count_RNA <- sum(rna_counts)
    Count_Ambient_RNA <- Total_Count_RNA - (sum(soupx_list[[obj]]))
    Total_Count_Corrected <- sum(soupx_list[[obj]])
     # Calculate ambient RNA percentage
    ambient_percentage <- (1 - (sum(soupx_list[[obj]]/sum(rna_counts))))*100
    ambient_percentage <- round(ambient_percentage, 1)
    print(paste0("ambiant RNA percentage for:", obj, "= ", ambient_percentage))
    result_row <- data.frame(Sample = obj, Total_Count_RNA = Total_Count_RNA ,  nCount_Ambient_RNA = Count_Ambient_RNA, nCount_corrected =Total_Count_RNA- Count_Ambient_RNA, Percentage_ambient = ambient_percentage, Percentage_correct = 100-ambient_percentage )
    results <- rbind(results, result_row)
    
  }
}

write.csv(results, file = paste0(fig_path_stnx, "01_02_Ambient_RNA_percentage.csv"), row.names = TRUE)
# results <- read.csv(paste0(fig_path_stnx, "01_02_Ambient_RNA_percentage.csv"), row.names = 1)

range(results$Percentage_ambient) # 1.0 44.9
mean(results$Percentage_ambient) # 7.744828
library(tidyr)
library(dplyr)
library(tidyverse)
# Reshape the data to long format
results_long <- results %>%
  pivot_longer(cols = c(nCount_corrected, nCount_Ambient_RNA), 
               names_to = "RNA_Type", 
               values_to = "Count")

# Create stacked bar plot with x-axis labels rotated 45 degrees
plot <- ggplot(results_long, aes(x = Sample, y = Count, fill = RNA_Type)) +
  geom_bar(stat = "identity") +
  labs(title = "Ambient RNA Count per Sample \n SoupX", x = "Sample", y = "RNA Count") +
  scale_fill_manual(values = c("nCount_corrected" = "steelblue", "nCount_Ambient_RNA" = "firebrick")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))





# Reorder the Sample factor by Total_Count_RNA from high to low
results$Sample <- fct_reorder(results$Sample, results$Total_Count_RNA, .desc = TRUE)

# Reshape the data to long format
results_long <- results %>%
  pivot_longer(cols = c(nCount_corrected, nCount_Ambient_RNA), 
               names_to = "RNA_Type", 
               values_to = "Count")

# Create stacked bar plot with x-axis labels rotated 45 degrees
plot2 <- ggplot(results_long, aes(x = Sample, y = Count, fill = RNA_Type)) +
  geom_bar(stat = "identity") +
  labs(title = "Ambient RNA Count per Sample \n SoupX", x = "Sample", y = "RNA Count") +
  scale_fill_manual(values = c("nCount_corrected" = "steelblue", "nCount_Ambient_RNA" = "firebrick")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Reorder the Sample factor by Total_Count_RNA from high to low
results$Sample <- fct_reorder(results$Sample, results$Percentage_ambient, .desc = TRUE)

# Reshape the data to long format
results_long <- results %>%
  pivot_longer(cols = c( Percentage_correct,Percentage_ambient), 
               names_to = "RNA_Type", 
               values_to = "Count")

# Create stacked bar plot with x-axis labels rotated 45 degrees
plot3 <- ggplot(results_long, aes(x = Sample, y = Count, fill = RNA_Type)) +
  geom_bar(stat = "identity") +
  labs(title = "Ambient RNA percentage per Sample \n SoupX", x = "Sample", y = "RNA Count") +
  scale_fill_manual(values = c("Percentage_correct" = "steelblue", "Percentage_ambient" = "firebrick")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Define the file path for the PDF output
pdf(file = paste0(fig_path_stnx, "01_02_Ambeint_RNA_stacked_bar_plot.pdf"), width = 30, height = 25, pointsize = 5)

# Print the plot
print(plot)
print(plot2)
print(plot3)
# Close the PDF device
dev.off()

```


#converting soupx mtx to seurat object
1- creating Seurat objects.
2- creating donor column

3- creating sample_id column

4- changing all barcodes  from to 1-38

5- merging all soupx seurat

6- changing all rownames to $barcodes

7- opening and adding the atac obj


```{r, soupx_seurat_list}
# Create a vector with numbers 1 through 25
AML_samples_numbers <- seq(1, 25)

# Additional numbers to add
Healthy_sample_numbers <- c(26, 29, 36, 37)

# Combine the two vectors
sample_number <- c(AML_samples_numbers, Healthy_sample_numbers)

# Update barcodes to include unique sample numbers which will be same as cellranger arc -aggr output

for (i in seq_along(soupx_list)) {
  numb <- sample_number[[i]]
  colnames(soupx_list[[i]]) <- sub("1$", numb, colnames(soupx_list[[i]]))
}

# Initialize list to store Seurat objects
soupx_seurat_list <- list()
# Create Seurat objects from SoupX-corrected counts
for (sobj in names(soupx_list)) {
  soupcount <- soupx_list[[sobj]]
  seurat_obj <- CreateSeuratObject(counts = soupcount)
  soupx_seurat_list[[sobj]] <- seurat_obj
}

# after an update the below code did not work on seurat object so i changed the barcodes in mtxes.
for (id in names(soupx_seurat_list)) {
  soupx_seurat_list[[id]]$donor <- id
}

# Check if barcode names match between SoupX-corrected and original Seurat objects
for (obj in names(seurat_list)) {
  if (identical(colnames(soupx_seurat_list[[obj]]), colnames(seurat_list[[obj]]))) {
    print(paste("Column names match for:", obj))
  } else {
    print(paste("Mismatch in column names for:", obj))
  }
}

# Check if gene names (row names) match between objects
for (obj in names(seurat_list)) {
  if (identical(rownames(soupx_seurat_list[[obj]]), rownames(seurat_list[[obj]]))) {
    print(paste("Row names match for:", obj))
  } else {
    print(paste("Mismatch in row names for:", obj))
  }
}


# Assign sample ID as cell identity (optional, useful for integration)
for (sample_id in names(soupx_seurat_list)) {
 Idents(soupx_seurat_list[[sample_id]]) <- sample_id
}
# Check if all Seurat objects have the same genes
all_rownames <- lapply(soupx_seurat_list, rownames)
are_identical <- length(unique(all_rownames)) == 1 #TRUE

# Check if all Seurat objects have the same genes
all_colnames <- lapply(soupx_seurat_list, colnames)
all_colnames_combined <- unlist(all_colnames, use.names = FALSE)

```


Creating corrected seurat list
#Merging all soupx seurat objects
in previous step defined a uniq cell barcode suffix for each sample, so change to add.cell.ids =NULL.
```{r}
packageVersion("scCustomize") #‘2.1.2’
Seurat_merge_soupx <- merge(x=soupx_seurat_list[[1]], y=soupx_seurat_list[2:length(soupx_seurat_list)], add.cell.ids = NULL,  project = "MiniAtlas")

identical(colnames(Seurat_merge_soupx),all_colnames_combined) #TRUE
Seurat_merge_soupx
# An object of class Seurat 
# 36601 features across 232173 samples within 1 assay 
# Active assay: RNA (36601 features, 0 variable features)
#  2 layers present: counts, data

saveRDS(Seurat_merge_soupx, paste0(rds_path_vst, "01_02_Seurat_merge_soupx_RNA_corrected_25_aml_4healthy.rds"))
```
#Reading ATAC data

The original ATAC data was generated using Cell Ranger ARC -aggr, meaning all ATAC samples share the same peak coordinates. 

For RNA data, we can directly load and convert it into a Seurat object. However, for ATAC data, we first need to convert it into a ChromatinAssay object.

Since I needed to remove ambient RNA on a per-sample basis, I used the Cell Ranger ARC count output for RNA data instead of the aggr output.

After filtering out ambient RNA from each sample, I merged all RNA samples and then integrated them with the ATAC data from the Cell Ranger ARC aggr output in the next step.



```{r,}
Seurat_merge_soupx <- readRDS(paste0(rds_path_vst, "01_02_Seurat_merge_soupx_RNA_corrected_25_aml_4healthy.rds"))
# output of sub-sampling of atac data in python saved in atac_dir dir.
atac_dir = '/stornext/General/data/user_managed/grpu_mritchie_1/reza/projects/mat_seq/RaCHseq_AML/cellranger-arc_outputs/aggr_outputs/AML_BMMC_38_Samples_no_secondary_analysis/subset_all_aml_atac_4healthy/'
library(Seurat)
# Expecting matrix.mtx.gzcd
atac_counts <- Seurat::Read10X(atac_dir) #184540 x 232173 sparse Matrix of class "dgCMatrix"

identical(colnames(Seurat_merge_soupx), colnames(atac_counts)) #FALSE #it means the order has changed in python, further investigation is needed in python.
atac_counts <- atac_counts[, colnames(Seurat_merge_soupx)]
identical(colnames(Seurat_merge_soupx), colnames(atac_counts)) #TRUE
library(EnsDb.Hsapiens.v86)

annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86) 
seqlevelsStyle(annotations) <- 'UCSC'
genome(annotations) <- "hg38"


grange_counts <- StringToGRanges(rownames(atac_counts), sep = c(":", "-"))
grange_use <- seqnames(grange_counts) %in% standardChromosomes(grange_counts)
atac_counts <- atac_counts[as.vector(grange_use), ] #184466 x 232173 sparse Matrix of class "dgCMatrix"

  # Read fragments
frag_file <-"/stornext/General/data/user_managed/grpu_mritchie_1/reza/projects/mat_seq/RaCHseq_AML/cellranger-arc_outputs/aggr_outputs/AML_BMMC_38_Samples_no_secondary_analysis/outs/atac_fragments.tsv.gz"

  # Create Chromatin Assay
  chrom_assay <- CreateChromatinAssay(
    counts = atac_counts,
    sep = c(":", "-"),
    genome = 'hg38',
    fragments = frag_file,
    min.cells = 10,
    annotation = annotations
  )


  


identical(colnames(chrom_assay), colnames(Seurat_merge_soupx)) #True

identical(colnames(chrom_assay), colnames(atac_counts)) #TRUE #it means the order has changed in python, further investigation is needed in python or randomly check with some some indvidual data from cellranger arc -count.




  # Add ATAC data to the Seurat object
Seurat_merge_soupx [["ATAC"]] <- chrom_assay
Seurat_merge_soupx

# An object of class Seurat 
# 221067 features across 232173 samples within 2 assays 
# Active assay: RNA (36601 features, 0 variable features)
#  2 layers present: counts, data
#  1 other assay present: ATAC


saveRDS(Seurat_merge_soupx, paste0(rds_path_vst, "01_02_Seurat_merge_Multiome_soupx_RNA_corrected_25_aml_4_healthy_232173_cells.rds"))
# Seurat_merge_soupx <- readRDS(paste0(rds_path_vst, "01_02_Seurat_merge_Multiome_soupx_RNA_corrected_25_aml_4_healthy_232173_cells.rds"))
```


